### 2018-04-21 9.0/paper.tex
Reshuffle some of the sections.

\documentclass{ansarticle-preprint}
%\usepackage{ucs}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
%\usepackage{cite}
\usepackage{anslistings}
\usepackage{multicol}

\usepackage{pgfplots}
\usepackage{pgfplotstable}

\usepackage{fontenc}
\usepackage{graphicx}

\pgfplotsset{compat=1.9}

\newcommand{\specialword}[1]{\texttt{#1}}
\newcommand{\dealii}{{\specialword{deal.II}}}
\newcommand{\pfrst}{{\specialword{p4est}}}
\newcommand{\trilinos}{{\specialword{Trilinos}}}
\newcommand{\aspect}{\specialword{Aspect}}
\newcommand{\petsc}{\specialword{PETSc}}
\newcommand{\cmake}{{\specialword{CMake}}}
\newcommand{\autoconf}{{\specialword{autoconf}}}

%
% Author list -- please add yourself in both places below (in
%                alphabetical order) if you think that your
%                contributions to the last release warrant this
%

\hypersetup{
  pdfauthor={
    Giovanni Alzetta,
    Daniel Arndt,
    Wolfgang Bangerth,
    Vishal Boddu,
    Benjamin Brands,
    Denis Davydov,
    Rene Gassm\"{o}ller,
    Timo Heister,
    Luca Heltai,
    Katharina Kormann,
    Martin Kronbichler,
    Matthias Maier,
    Jean-Paul Pelteret,
    Bruno Turcksin,
    David Wells
  },
  pdftitle={The deal.II Library, Version 9.0, 2018},
}

\title{The \dealii{} Library, Version 9.0}

\author[1]{Giovanni Alzetta}
\affil[1]{SISSA,
  International School for Advanced Studies,
  Via Bonomea 265,
  34136, Trieste, Italy.
{\texttt{galzetta@sissa.it}}}

\author[2]{Daniel Arndt}
\affil[2]{Interdisciplinary Center for Scientific Computing,
  Heidelberg University,
  Im Neuenheimer Feld 205,
  69120 Heidelberg, Germany.
  {\texttt{daniel.arndt@iwr.uni-heidelberg.de}}}

\author[3]{Wolfgang Bangerth}
\affil[3]{Department of Mathematics, Colorado State University, Fort
  Collins, CO 80523-1874, USA.
    {\texttt{bangerth@colostate.edu}}}

\author[4]{Vishal Boddu}
\affil[4]{Chair of Applied Mechanics,
  Friedrich-Alexander-Universit\"{a}t Erlangen-N\"{u}rnberg,
  Egerlandstr.\ 5,
  91058 Erlangen, Germany.
  {\texttt{vishal.boddu@fau.de}}}

\author[5]{Benjamin Brands}
\affil[5]{Chair of Applied Mechanics,
          Friedrich-Alexander-Universit\"{a}t Erlangen-N\"{u}rnberg,
          Egerlandstr.\ 5, 91058 Erlangen, Germany.
          {\texttt{benjamin.brands@fau.de}}}

\author[6]{Denis Davydov}
\affil[6]{Chair of Applied Mechanics, Friedrich-Alexander-Universit\"{a}t Erlangen-N\"{u}rnberg,
            Egerlandstr.\ 5, 91058 Erlangen, Germany.
    {\texttt{denis.davydov@fau.de}}}

\author[7]{Rene Gassm\"{o}ller}
\affil[7]{Department of Earth and Planetary Sciences,
  University of California Davis,
  One Shields Avenue,
  CA-95616 Davis, USA.
  {\texttt{rgassmoeller@ucdavis.edu}}}

\author[8]{Timo Heister}
\affil[8]{Mathematical Sciences,
  O-110 Martin Hall,
  Clemson University,
  Clemson, SC 29634, USA.
  {\texttt{heister@clemson.edu}}}
%
\author[9]{Luca Heltai}
\affil[9]{SISSA,
  International School for Advanced Studies,
  Via Bonomea 265,
  34136, Trieste, Italy.
{\texttt{luca.heltai@sissa.it}}}

\author[10]{Katharina Kormann}
\affil[10]{Max Planck Institute for Plasma Physics,
Boltzmannstr.~2, 85748 Garching, Germany.
{\texttt{katharina.kormann@ipp.mpg.de}}}

\author[11]{Martin Kronbichler}
\affil[11]{Institute for Computational Mechanics,
  Technical University of Munich,
  Boltzmannstr.~15, 85748 Garching, Germany.
  {\texttt{kronbichler@lnm.mw.tum.de}}}

\author[12]{Matthias Maier}
\affil[12]{School of Mathematics,
  University of Minnesota,
  127 Vincent Hall, 206 Church Street SE,
  Minneapolis, MN 55455, USA.
  {\texttt{msmaier@umn.edu}}}

\author[13]{Jean-Paul Pelteret}
\affil[13]{Chair of Applied Mechanics,
  Friedrich-Alexander-Universit\"{a}t Erlangen-N\"{u}rnberg,
  Egerlandstr.\ 5,
  91058 Erlangen,
  Germany.
  {\texttt{jean-paul.pelteret@fau.de}}}

 \author[14]{Bruno Turcksin\footnote{
   This manuscript has been authored by UT-Battelle, LLC under Contract No.
   DE-AC05-00OR22725 with the U.S. Department of Energy. The United States
   Government retains and the publisher, by accepting the article for
   publication, acknowledges that the United States Government retains a
   non-exclusive, paid-up, irrevocable, worldwide license to publish or reproduce
   the published form of this manuscript, or allow others to do so, for United
   States Government purposes. The Department of Energy will provide public
   access to these results of federally sponsored research in accordance with the
   DOE Public Access Plan (http://energy.gov/downloads/doe-public-access-plan).}}
 \affil[14]{Computational Engineering and Energy Sciences Group,
   Computional Sciences and Engineering Division,
   Oak Ridge National Laboratory, 1 Bethel Valley Rd.,
   TN 37831, USA.
   {\texttt{turcksinbr@ornl.gov}}}

\author[15]{David Wells}
\affil[15]{Department of Mathematical Sciences, Rensselaer Polytechnic
Institute, Troy, NY 12180, USA.
  {\texttt{wellsd2@rpi.edu}}}

\renewcommand{\labelitemi}{--}

\begin{document}
\maketitle

\begin{abstract}
  This paper provides an overview of the new features of the finite element
  library \dealii{} version 9.0.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Overview}

\dealii{} version 9.0.0 was released April X, 2018.
\marginpar{Update the date}
This paper provides an
overview of the new features of this major release and serves as a citable
reference for the \dealii{} software library version 9.0. \dealii{} is an
object-oriented finite element library used around the world in the
development of finite element solvers. It is available for free under the
GNU Lesser General Public License (LGPL) from the \dealii{} homepage at
\url{http://www.dealii.org/}.

The major changes of this release are:
\begin{itemize}
\item
  Manifold smoothing: The manifold smoothing algorithms applied in the
  Triangulation class and MappingQGeneric have been changed from the old
  Laplace-style smoothing to a transfinite interpolation that linearly
  blends between the descriptions on the faces around a cell. The old
  transformation introduced boundary layers inside cells that prevented
  convergence rates from exceeding \(3.5\) in the global \(L^2\) errors on typical
  settings. This change also considerably improves mesh quality on settings
  where curved descriptions are only applied to the boundary rather than
  the whole volume.

\item
  A dedicated particles module has been created to provide support for the
  storage of particles and their properties. This module provides a base class
  \texttt{Particles} which represents a particle with position, an ID number and
  a variable number of properties. It also provides a \texttt{ParticleHandler}
  class that manages the storage and handling of particles. This class
  distributes the particles in distributed domains (these are the only domains
  supported by this class so far)

\item
  A dedicated differentiation module has been created to provide support for
  automatic and symbolic differentiation. Automatic differentiation is a set of
  technique to evaluate the derivatives of a function defined by a compute
  program. Currently the Adol-C and Sacado libraries are supported through a
  unified interface. In practice, this support means that Adol-c and Sacado
  data types can be used in the \texttt{FEValues} and \texttt{FEValuesViews}
  to represent degree-of-freedom values.

\item Interfaces to more external libraries and programs.

\item C++11 is now both required and used.

\item
  Support for GPU computations.

\item \dealii{} has made extensive use of both the Clang-Tidy \cite{clang-tidy}
  and Coverity Scan \cite{coverity} static analysis tools for detecting bugs
  and other issues in the code. For example, around 260 issues were detected and
  fixed using the latter tool.
      %% > Is there somewhere we can cite for either of these tools? It would
      %% > be nice to recognize the department of homeland security for offering
      %% The contract with homeland security was completed in 2009.

\item
  \texttt{LinearOperator}, a flexible template class that implements the action of a
  linear operator (see \cite{MaierBardelloniHeltai-2016-b}), now supports
  computations with Trilinos, Schur complements, and linear constraints. This
  class is, as of this release, the official replacement for about half a dozen
  similar (but less general) classes, such as \texttt{FilteredMatrix},
  \texttt{IterativeInverse}, and \texttt{PointerMatrix}.

\item
  Significant extension of matrix-free capabilities, including support for face
  integrals in discontinuous Galerkin schemes, a new Hermite-like polynomial
  basis adapted to face integrals that involve derivatives of shape functions,
  and several performance enhancements.

\item New non-standard quadrature rules:
A number of non-standard, special-purpose quadrature rules have been
implemented. Among these are ones for
(i) truncating standard formulas to simplical domains (\texttt{Qsimplex});
(ii) singular transformations of the unit cell to the unit simplex
    (\texttt{QDuffy});
(iii) composition of simplical quadrature rules to a combined rule on the
    unit cell (\texttt{QSplit});
and (iv) transformation of the unit square to polar coordinates
    (\texttt{QTrianglePolar}).
These quadrature rules greatly help when integrating singular
functions or on singular domains. They
are mainly used in Boundary Element Methods.

\item
  Support for complex-valued vectors at the same level as real-valued
  vectors.

\item New python tutorial program tutorial-1; as well as
  updates to step-37. In addition, the separate code
  gallery of \dealii{} has gained a number of new entries.

  \item More than 330 other features and bugfixes.
\end{itemize}
The more important ones of these changes will be detailed in the
following section.  Information on how to cite \dealii{} is provided
in Section \ref{sec:cite}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Significant changes to the library}

This release of \dealii{} contains a number of large and significant changes
that will be discussed in the following sections. It of course also contains a
vast number of smaller changes and added functionality; the details of these
can be found
\href{https://www.dealii.org/developer/doxygen/deal.II/changes_between_8_5_and_9_0.html}{
in the file that lists all changes for this release}, see \cite{changes90}.
(The file is also linked to from the web site of each release as well as
the release announcement.)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Manifolds}

...

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Support for particle-in-cell methods}

...

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Dedicated support for automatic and symbolic differentiation}

...

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{New interfaces to external libraries and programs}

\dealii{} has always tried to leverage high-quality implementations of
algorithms available through other open source software, rather than
re-implementing their functionality. (A list of interfaces to other
packages is given in Section~\ref{sec:cite}.) As part of the current
release, we have written several new interfaces as discussed in the following.

\marginpar{Add references to all of these libraries.}

\paragraph*{ScaLAPACK, a parallel dense linear algebra library.}
  ScaLAPACK provides block-cyclic matrix distribution over 2D process
  grid. The functionality and interface of our wrappers is similar to
  the LAPACK wrappers for serial dense linear algebra, namely
  matrix-matrix multiplication, Cholesky and LU factorizations,
  eigensolvers, SVD, pseudoinverses, and save/load operations using
  the HDF5 format. All of this functionality is available even in
  cases where the number of MPI processes does not exactly divide the
  number of tiles stored in the block-cyclic data distribution.

  As part of this effort, we have also improved LAPACK support: there
  are now methods to perform rank-1 updates/downdates, Cholesky
  factorizations, to compute the trace and determinant, as well as
  estimate the reciprocal condition number. We also now support
  configuration with 64-bit BLAS.

\paragraph*{SUNDIALS, a SUite of Nonlinear and DIfferential/ALgebraic
  Equation Solvers.}
  Solving nonlinear algebraic and differential equations is both a
  common task and one that often requires sophisticated globalization
  algorithms for efficiency and reliability. SUNDIALS provides these
  in a widely used format, both sequentially and in parallel.

  \dealii{} now has interfaces to SUNDIALS's ARKode, IDA, and KINSOL sub-packages.
  ARKode is a solver library that provides adaptive-step time
  integration. IDA is a package for the solution of differential-algebraic
  equations systems in the form $F(t,y,y')=0$. KINSOL is solver for nonlinear
  algebraic systems.

\paragraph*{ROL, a Rapid Optimization Library.}
ROL is a package for large-scale optimization.
\marginpar{Can we add a sentence or two about the ROL interface here?}

\paragraph*{nanoflann, a library for building and querying
  $k$-d trees of datasets.} Operations such as finding the vertex
or cell closest
to a given evaluation point occur frequently in many applications that
use unstructured meshes. While the naive algorithm is linear in the
number of vertices or cells, many such operations can be made
significantly faster by building a $k$-d tree data structure that
recursively subdivides a $k$ dimensional space. The nanoflann library
provides such a data structure and allows querying it, either for
closest points (e.g., when finding the closest vertex) or for
searching the points that fall within a radius of a target point. This
functionality is now available via \dealii{} interfaces.

\paragraph*{Assimp, the Open Asset Import Library.}
  Assimp can be used to read about 40 different 3D graphics formats. A
  subset of these formats can be now be read from within \dealii{} to
  generate two-dimensional meshes, possibly embedded in a
  three-dimensional space.

\paragraph*{Gmsh.}
  \marginpar{Do we actually provide any functionality with the GMSH
  executable yet? If so, complete this part}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Use of C++11}
\dealii{} first offered support for a subset of C++11 features in
version 6.2, released in 2009. The current release is the first to
\emph{require} a C++11 compiler.

Many parts of the code base have been rewritten to both support and
use the new features of C++11. In particular, \dealii{} now makes
extensive use of move semantics as well as range-based \texttt{for}
loops with \texttt{auto} type deduction of iterator variables. We have
also largely replaced \texttt{push\_back()} by
\texttt{emplace\_back()} when adding elements to collections more
efficiently.

Finally, we have changed the entire code base to avoid using raw
pointers and instead use \texttt{std::unique\_ptr} and
\texttt{std::shared\_ptr} where possible to make memory management
more reliable. These changes include some minor incompatibilities: all
\texttt{clone()} functions (such as \texttt{FiniteElement::clone()} and
\texttt{Mapping::clone()}) now return \texttt{std::unique\_ptr}s instead of
C-style raw pointers. Indeed, nearly all interfaces throughout the library that return a pointer
now return either a \texttt{std::shared\_ptr} or a \texttt{std::unique\_ptr},
thereby clarifying object ownership responsibilities and avoiding memory leaks.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Support for GPU computations}

\marginpar{Bruno?}
...

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{New and updated tutorial programs}

\marginpar{OLD}
In addition to the updated tutorial programs mentioned in the previous
section, this release of \dealii{} includes three new tutorials TODO: old:
\begin{itemize}
 \item {\bf step-55} explains how to solve the Stokes
 equations efficiently in parallel. It is a good introduction to solving
 systems of PDEs in parallel, discusses optimal block
 preconditioners, and demonstrates other aspects like error computation.
 Inverses of individual blocks of the linear system are approximated with an algebraic
 multigrid preconditioner.

 \item {\bf step-56} shows how to apply geometric multigrid
 preconditioners on a subset of a system of PDEs. The problem solved here is the
 Stokes equations, like in step-55.

 \item {\bf step-57} solves the stationary Navier-Stokes equations.
 The nonlinear system is solved using Newton's method on a sequence of adaptively refined
 grids. The preconditioner is again built on a block factorization of the saddle point
 system like in step-55 and step-56, but the non-symmetric terms stemming from the
 nonlinear convective part requires more sophisticated solvers. The benchmark problem,
 flow in the 2d lid-driven cavity, requires a continuation method for high Reynolds
 numbers.
 \end{itemize}

In addition to tutorials, \dealii{} has a separate ``code gallery'' that
consists of programs shared by users as examples of what can be
done with \dealii{}. While not part of the release process, it is nonetheless
worth mentioning that the set of new programs since the last release covers
the following topics:
\marginpar{OLD}
  \begin{itemize}
    \item Quasi-static quasi-incompressible visco-elastic material behavior;
    \item Multiphase Navier-Stokes flow;
    \item The evolution of global-scale topography on planetary bodies;
    \item Goal-oriented elastoplasticity.
  \end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Incompatible changes}

\marginpar{OLD}
The 9.0 release includes around XYZ
\href{https://www.dealii.org/developer/doxygen/deal.II/changes_between_8_5_and_9_0.html}{
incompatible changes}; see \cite{changes90}. The majority of these changes
should not be visible to typical user codes; some remove previously
deprecated classes and functions, and the majority change internal
interfaces that are not usually used in external applications. However, three
incompatible changes are worth mentioning:
\begin{itemize}
  \item High-order Lagrange elements, both continuous \verb!FE_Q! and
    discontinuous \verb!FE_DGQ! types, now use the nodal points of the
    Gauss-Lobatto quadrature formula as support points by default, rather than the
    previous equidistant ones. For cubic polynomials and higher, the point
    distribution has thus changed and, consequently, the entries in
    solution vectors will be different compared to previous
    versions of \dealii{}. Note, however, that using the Gauss-Lobatto points as nodal
    points results in a much more stable interpolation, including better
    iteration counts in most iterative solvers.
  \item
    The library no longer instantiates template classes with \texttt{long
    double}. These were rarely used, but took up a significant
    fraction of compile and link time, as well as library
    size. Application programs can, however, still instantiate all
    template classes with \texttt{long
    double} as long as they include the corresponding \texttt{.templates.h}
    header files.
  \item
    The \texttt{ParameterGUI} has been moved to a separate repository.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{How to cite \dealii{}}\label{sec:cite}

In order to justify the work the developers of \dealii{} put into this
software, we ask that papers using the library reference one of the
\dealii{} papers. This helps us justify the effort we put into it.

There are various ways to reference \dealii{}. To acknowledge the use of
the current version of the library, \textbf{please reference the present
document}. For up to date information and bibtex snippets for this document
see:
\begin{center}
 \url{https://www.dealii.org/publications.html}
\end{center}

The original \texttt{\dealii{}} paper containing an overview of its
architecture is \cite{BangerthHartmannKanschat2007}. If you rely on
specific features of the library, please consider citing any of the
following:
\begin{itemize}
 \item For geometric multigrid: \cite{Kanschat2004,JanssenKanschat2011};
 \item For distributed parallel computing: \cite{BangerthBursteddeHeisterKronbichler11};
 \item For $hp$ adaptivity: \cite{BangerthKayserHerold2007};
  \item For $PUM$ and enrichment of the FE space: \cite{Davydov2016};
 \item For matrix-free and fast assembly techniques:
   \cite{KronbichlerKormann2012};
 \item For computations on lower-dimensional manifolds:
   \cite{DeSimoneHeltaiManigrasso2009};
 \item For integration with CAD files and tools:
   \cite{HeltaiMola2015};
 \item For \texttt{LinearOperator} and \texttt{PackagedOperation} facilities:
   \cite{MaierBardelloniHeltai-2016-a,MaierBardelloniHeltai-2016-b}.
 \item For uses of the \texttt{WorkStream} interface:
   \cite{TKB16}.
\end{itemize}

\marginpar{update}
\dealii{} can interface with many other libraries:
\begin{multicols}{3}
\begin{itemize}
\item Adol-c \cite{adol-c}
\item ARPACK \cite{arpack}
\item Assimp \cite{assimp}
\item BLAS, LAPACK
\item GSL \cite{gsl2016}
\item HDF5 \cite{hdf5}
\item METIS \cite{karypis1998fast}
\item MUMPS \cite{ADE00,MUMPS:1,MUMPS:2,mumps-web-page}
\item muparser \cite{muparser-web-page}
\item nanoflann \cite{nanoflann}
\item NetCDF \cite{rew1990netcdf}
\item OpenCASCADE \cite{opencascade-web-page}
\item p4est \cite{p4est}
\item PETSc \cite{petsc-user-ref,petsc-web-page}
\item SLEPc \cite{Hernandez:2005:SSF}
\item SUNDIALS \cite{sundials}
\item TBB \cite{Rei07}
\item Trilinos \cite{trilinos,trilinos-web-page}
\item UMFPACK \cite{umfpack}
\end{itemize}
\end{multicols}
Please consider citing the appropriate references if you use interfaces to these
libraries.

Older releases of \dealii{} can be cited as
\cite{dealII80,dealII81,dealII82,dealII83,dealII84,dealII85}.

\nocite{BangerthKanschat1999}

\section{Acknowledgments}

\dealii{} is a world-wide project with dozens of contributors around the
globe. Other than the authors of this paper, the following people contributed code to
this release:\\
%
% get this from the changes/*/* files using the command listed in the
% release-tasks paper and remove the authors of this paper
%
% Updated 2018-04-19
Julian Andrej,
Rajat Arora,
Lucas Campos,
Praveen Chandrashekar,
Jie Cheng,
Emma Cinatl,
Conrad Clevenger,
Ester Comellas,
Sambit Das,
Giovanni Di Ilio,
Nivesh Dommaraju,
Marc Fehling,
Niklas Fehn,
Menno Fraters,
Anian Fuchs,
Daniel Garcia-Sanchez,
Nicola Giuliani,
Anne Glerum,
Christoph Goering,
Alexander Grayver,
Samuel Imfeld,
Daniel Jodlbauer,
Guido Kanschat,
Vishal Kenchan,
Andreas Kergassner,
Eldar Khattatov,
Ingo Kligge,
Uwe Köcher,
Joachim Kopp,
Ross Kynch,
Konstantin Ladutenko,
Tulio Ligneul,
Karl Ljungkvist,
Santiago Ospina,
Alexey Ozeritsky,
Dirk Peschka,
Simon Puchert,
E. G. Puckett,
Lei Qiao,
Ce Qin,
Jonathan Robey,
Alberto Sartori,
Daniel Shapero,
Ben Shields,
Simon Sticko,
Oliver Sutton,
Zhuoran Wang,
Xiaoyu Wei,
Michał Wichrowski,
Julius Witte,
Feimi Yu,
Weixiong Zheng.

Their contributions are much appreciated!

\bigskip

\dealii{} and its developers are financially supported through a
variety of funding sources:
\marginpar{update}

D.~Arndt was supported by the German Research Foundation (DFG) under the
project ``High-order discontinuous Galerkin for the exa-scale'' (ExaDG) within the
priority program ``Software for Exascale Computing'' (SPPEXA).

W.~Bangerth and R.~Gassm\"{o}ller were partially
supported by the National Science Foundation under award OCI-1148116
as part of the Software Infrastructure for Sustained Innovation (SI2)
program; and by the Computational Infrastructure in Geodynamics initiative
(CIG), through the National Science Foundation under Awards
No.~EAR-0949446 and EAR-1550901 and The University of California -- Davis.

V.~Boddu was supported by the German Research Foundation (DFG) under the
research group project FOR 1509.

B.~Brands was partially supported by the Indo-German exchange programm
``Multiscale Modeling, Simulation and Optimization for Energy, Advanced
Materials and Manufacturing'' (MMSO) funded by DAAD (Germany) and UGC
(India).

D.~Davydov was supported by the German Research Foundation (DFG), grant DA
1664/2-1.

T.~Heister was partially supported by the Computational Infrastructure in
Geodynamics initiative (CIG), through the National Science Foundation (NSF)
under Award No. EAR-0949446 and EAR-1550901 and The University of
California -- Davis, and NSF Grant DMS-1522191.

M.~Kronbichler was partially supported by the German Research Foundation (DFG)
under the project ``High-order discontinuous Galerkin for the exa-scale''
(ExaDG) within the priority program ``Software for Exascale Computing''
(SPPEXA), grant agreement no.~KR4661/2-1 and the Bayerisches Kompetenznetzwerk
f\"ur Technisch-Wissenschaftliches Hoch- und H\"ochstleistungsrechnen
(KONWIHR).

M.~Maier was partially supported by ARO MURI Award No. W911NF-14-0247.

J-P.~Pelteret was supported by the European Research Council (ERC) through
the Advanced Grant 289049 MOCOPOLY.

B.~Turcksin: This material is based upon work supported by the U.S.
Department of Energy, Office of Science, under contract number
DE-AC05-00OR22725.

D.~Wells was supported by the National Science Foundation (NSF) through Grant
DMS-1344962.

The Interdisciplinary Center for Scientific Computing (IWR) at Heidelberg
University has provided hosting services for the \dealii{} web page.

\bibliography{paper}{}
\bibliographystyle{abbrv}

\end{document}