### 2019-05-22 9.1/paper.tex
Replace cores -> MPI ranks.
I will run on 300k cores only next week, up to now it was only 152k cores
and 304k MPI ranks

\documentclass{ansarticle-preprint}
%\usepackage{ucs}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
%\usepackage{cite}
\usepackage{anslistings}
\usepackage{multicol}
\usepackage{pdfsync}

\usepackage{pgfplots}
\usepackage{pgfplotstable}

\usepackage{fontenc}
\usepackage{graphicx}
\usepackage{xspace}

\pgfplotsset{compat=1.9}

\newcommand{\specialword}[1]{\texttt{#1}}
\newcommand{\dealii}{{\specialword{deal.II}}\xspace}
\newcommand{\pfrst}{{\specialword{p4est}}\xspace}
\newcommand{\trilinos}{{\specialword{Trilinos}}\xspace}
\newcommand{\aspect}{\specialword{Aspect}\xspace}
\newcommand{\petsc}{\specialword{PETSc}\xspace}
\newcommand{\cmake}{{\specialword{CMake}}\xspace}
\newcommand{\autoconf}{{\specialword{autoconf}}\xspace}

%
% Author list -- please add yourself in both places below (in
%                alphabetical order) if you think that your
%                contributions to the last release warrant this
%

\hypersetup{
  pdfauthor={
    Daniel Arndt,
    Wolfgang Bangerth,
    Denis Davydov,
    Timo Heister,
    Luca Heltai,
    Martin Kronbichler,
    Matthias Maier,
    Jean-Paul Pelteret,
    Bruno Turcksin,
    David Wells
  },
  pdftitle={The deal.II Library, Version 9.1, 2019},
}

\title{The \dealii Library, Version 9.1}

\author[*1]{Daniel Arndt}
\affil[1]{Computational Engineering and Energy Sciences Group,
  Computional Sciences and Engineering Division,
  Oak Ridge National Laboratory, 1 Bethel Valley Rd.,
  TN 37831, USA.
  {\texttt{\{arndtd,turcksinbr\}@ornl.gov}}}

\author[2]{Wolfgang~Bangerth}
\affil[2]{Department of Mathematics, Colorado State University, Fort
  Collins, CO 80523-1874, USA.
  {\texttt{bangerth@colostate.edu}}}

\author[3]{Denis~Davydov}
\affil[3]{Chair of Applied Mechanics,
  Friedrich-Alexander-Universit\"{a}t Erlangen-N\"{u}rnberg,
  Egerlandstr.\ 5,
  91058 Erlangen, Germany.
  {\texttt{\{denis.davydov,jean-paul.pelteret\}@fau.de}}}

\author[4]{Timo~Heister}
\affil[4]{Mathematical Sciences,
  O-110 Martin Hall,
  Clemson University,
  Clemson, SC 29634, USA
  and Scientific Computing and Imaging Institute,
  72 S Central Campus Drive, Room 3750
  Salt Lake City, UT 84112.
  {\texttt{heister@sci.utah.edu}}}
%
\author[5]{Luca~Heltai}
\affil[5]{SISSA,
  International School for Advanced Studies,
  Via Bonomea 265,
  34136, Trieste, Italy.
  {\texttt{luca.heltai@sissa.it}}}

\author[6]{Martin~Kronbichler}
\affil[6]{Institute for Computational Mechanics,
  Technical University of Munich,
  Boltzmannstr.~15, 85748 Garching, Germany.
  {\texttt{kronbichler@lnm.mw.tum.de}}}

\author[7]{Matthias~Maier}
\affil[7]{Department of Mathematics,
  Texas A\&M University,
  3368 TAMU,
  College Station, TX 77845, USA.
  {\texttt{maier@math.tamu.edu}}}

\author[4]{Jean-Paul~Pelteret}

\author[*1]{Bruno~Turcksin}

\author[9]{David Wells}
\affil[9]{Department of Mathematics, University of North Carolina,
  Chapel Hill, NC 27516, USA.
  {\texttt{drwells@email.unc.edu}}}

\renewcommand{\labelitemi}{--}

\begin{document}
\maketitle

\footnotetext{%
  $^\ast$ This manuscript has been authored by UT-Battelle, LLC under Contract No.
  DE-AC05-00OR22725 with the U.S. Department of Energy. The United States
  Government retains and the publisher, by accepting the article for
  publication, acknowledges that the United States Government retains a
  non-exclusive, paid-up, irrevocable, worldwide license to publish or reproduce
  the published form of this manuscript, or allow others to do so, for United
  States Government purposes. The Department of Energy will provide public
  access to these results of federally sponsored research in accordance with the
  DOE Public Access Plan (http://energy.gov/downloads/doe-public-access-plan).}

\begin{abstract}
  This paper provides an overview of the new features of the finite element
  library \dealii version 9.1.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Overview}

\dealii version 9.1.0 was released May XX, 2019
This paper provides an
overview of the new features of this major release and serves as a citable
reference for the \dealii software library version 9.1. \dealii is an
object-oriented finite element library used around the world in the
development of finite element solvers. It is available for free under the
GNU Lesser General Public License (LGPL). Downloads are available at
\url{https://www.dealii.org/} and \url{https://github.com/dealii/dealii}.

The major changes of this release are:
%
\begin{itemize}
  \item Improved support for automatic and symbolic differentiation (see
    Section~\ref{subsec:ad}),
  \item Full support for $hp$ adaptivity in parallel computations (see
    Section~\ref{subsec:hp}),
  \item Interfaces to the HDF5 file format and libraries (see
    Section~\ref{subsec:hdf5}),
  \item Significantly extended GPU support (see Section~\ref{subsec:gpu}),
  \item Parallel geometric multigrid (GMG) improvements (see
    \cite{ClevengerHeisterKanschatKronbichler2019} and
    Section~\ref{subsec:gmg}),
  \item Four new tutorial programs (step-61, step-62, step-63, step-64),
    as well as one new code gallery program (see
    Section~\ref{subsec:steps}).
\end{itemize}
%
The major changes are discussed in detail in Section~\ref{sec:major}. There
are a number of other noteworthy changes in the current \dealii{} release
that we briefly outline in the remainder of this section:
%
\begin{itemize}
\item
  The release contains bug fixes and performance improvements of the
  matrix-free framework and related geometric multigrid solvers. In
  particular, the implementation of the Chebyshev iteration, an often used
  smoother in the matrix-free context, has been revised to reduce vector
  accesses. This speeds up matrix-free multigrid solvers by up to 10-15\%
  on geometries with affine (parallelogram and parallelpiped) cells, and up
  to 5\% on geometries with cells bounded by curved edges and faces.

\item
  Various variants of geometric multigrid solvers and matrix-free
  implementations were run on up to 304,128 MPI ranks during the acceptance
  phase of the SuperMUC-NG supercomputer, verifying the scalability of our
  implementations. Some performance bottlenecks on distributed
  triangulation with more than 100k ranks have been fixed.

\item A new class \texttt{ParsedConvergenceTable} has been introduced
  that greatly simplifies the construction of convergence tables,
  reading the options for the generation of the table from a parameter
  file, and providing methods that, combined with a parameter file,
  allow one to generate convergence tables using one-liners in user
  codes.

\item
  The \texttt{FE\_BernardiRaugel} class implements the non-standard
  Bernardi-Raugel (BR) element that can be used to construct a stable
  velocity-pressure pair for the Stokes equation \cite{BR85}. The BR
  element is an enriched version of the $Q_1^d$ element with added bubble
  functions on each edge (in 2d) or face (in 3d). It addresses the fact
  that the $Q_1^d\times Q_0$ combination is not inf-sup stable (requiring a
  larger velocity space), and that the $Q_2^d\times Q_1$ combination is
  stable but sub-optimal since the velocity space is too large relative to
  the pressure space to provide additional accuracy commensurate with the
  cost of the large number of velocity unknowns. The BR space is
  intermediate to the $Q_1^d$ and $Q_2^d$ spaces.

  The element is currently only implemented for parallelogram meshes due to
  difficulties associated with the mapping of shape functions: The shape
  functions of the $Q_1^d$ part of the element need to be mapped as
  scalars, as is common for the vector components of the $Q_1^d$ element;
  on the other hand, the vector-valued edge bubble functions need to be
  mapped using the Piola transform as is common for the Raviart-Thomas
  element. \dealii{} does not currently have the ability to use different
  mappings for individual shape functions, though this functionality is
  planned for the next release.

\item
  The \texttt{FE\_NedelecSZ} class is a new implementation of the
  N{\'e}d{\'e}lec element on quadrilaterals and hexahedra. It is based on
  the work of Zaglmayr \cite{Zag06} and overcomes the sign conflict issues
  present in traditional N{\'e}d{\'e}lec elements that arise from the edge
  and face parameterizations used in the basis functions. Therefore, this
  element should provide consistent results for general quadrilateral and
  hexahedral elements for which the relative orientations of edges and
  faces (as seen from all adjacent cells) are often difficult to establish.
  The \texttt{FE\_NedelecSZ} element addresses the sign conflict problem by
  assigning a globally defined orientation to local edges and faces. A
  detailed overview of the implementation of the \texttt{FE\_NedelecSZ}
  element in \dealii{} can be found in \cite{Kynch2017}.

\item All of the elementary geometrical objects of the library (namely
  \texttt{Point<dim>}, \texttt{Segment<dim>}, and
  \texttt{BoundingBox<dim>}) have been augmented with the traits
  needed to comply with \texttt{boost::geometry} concepts. A new
  interface to \texttt{boost::geometry::index::rtree} has been added
  that simplifies the construction of spacial indices based on points,
  bounding boxes, or segments.
\end{itemize}
%
In addition to these changes, the changelog lists more than 200 other
features and bugfixes.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Major changes to the library}
\label{sec:major}

This release of \dealii contains a number of large and significant changes
that will be discussed in this section.

It of course also contains a
vast number of smaller changes and added functionality; the details of these
can be found
\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_0_1_and_9_1_0.html}{
in the file that lists all changes for this release}, see \cite{changes91}.
(The file is also linked to from the web site of each release as well as
the release announcement.)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Improved support for automatic and symbolic differentiation}
\label{subsec:ad}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Full support for $hp$ adaptivity in parallel computations}
\label{subsec:hp}

\dealii{} has had support for $hp$ adaptive methods since around 2005
(documented in \cite{BangerthKayserHerold2007}) and for parallel
computations on distributed meshes since around 2010 (see
\cite{BangerthBursteddeHeisterKronbichler11}), but not for both at the
same time. The challenges to combine these are related to a number of
areas:
\begin{enumerate}
\item Data structures: The data structures necessary to store the
  indices of degrees of freedom are necessarily substantially more
  complicated for $hp$ algorithms than for the $h$-adaptive schemes
  that were already implemented. This is because the number of degrees
  of freedom per cell is now no longer constant. Furthermore,
  faces and edges may need to store more than one set of indices if
  the adjacent cells use different polynomial degrees; in the case of
  edges, the number of sets of indices may also be of variable size.

  All of this poses challenges in the parallel context because some of
  the information may not be known, or not be known right away, for
  cells that are not locally owned (i.e., for ghost and ``artificial''
  cells), and for which the data structures stored on different
  processors have to reconciled.

\item Algorithms: Already for $h$ adaptive meshes, enumerating all
  degrees of freedom on the global mesh is difficult as evidenced by
  the complications of the algorithms shown in Section 3.1 of
  \cite{BangerthBursteddeHeisterKronbichler11}, which requires more
  than a page of text and is implemented in many hundreds of lines of
  code).

  These difficulties are even more pronounced when using $hp$
  adaptivity. The main obstacle is the desire to unify the indices of
  matching degrees of freedom on adjacent cells, such as the edge
  degree of freedom of a $Q_2$ element with the middle one of
  the three edge degrees of freedom of a $Q_4$ element on a
  neighboring cell. Section 4.2 of
  \cite{BangerthKayserHerold2007} discusses a sequential algorithm
  that eliminates one of these degrees of freedom in favor of another,
  but in this introduces a ``master'' and a ``slave'' side of the
  interface. This is of no major consequence in sequential
  computations, but is inconvenient in parallel computations if the
  ``master'' side is a ghost cell whose degree of freedom indices are
  not (yet) available while enumerating local degrees of freedom, or
  if the master is an artificial cell whose information will never be
  available on a processor.

  An earlier implementation of the algorithm enumerating degrees of
  freedom, already available in \dealii{} 9.0, simply did not unify
  indices on processor boundaries. However, this makes the total
  number of degrees of freedom dependent on both the partition of the
  mesh and the number of processors available. We have therefore
  re-implemented the algorithm so that the unification does happen
  also on processor boundaries, and will report on the details
  elsewhere.

\item Data transfer patterns: An important algorithm in parallel
  finite element methods is the exchange of information stored on
  cells during mesh repartitioning. This happens, for example, when
  interpolating the solution from one mesh to the next,
  adaptively-refined mesh, or when adapting the polynomial degrees
  associated with each cell and repartitioning in order to
  balance the computational cost of each processor's partition. When
  using $h$ adaptive methods, the amount of data associated with each
  cell is fixed and the algorithms that implement the data transfer
  are consequently relatively simple. On the other hand, in $hp$
  contexts, each cell may have a different number of unknowns
  associated with it, and the algorithms that transfer the data are
  substantially more complicated. Furthermore, the amount of data
  associated with each cell may be large on cells with higher
  polynomial degrees, and might profit from compression before
  sending.

\item Balancing computational cost: For $h$-adaptive algorithms, the
  amount of work associated with each cell is essentially the same
  both during the assembly of linear systems as well as during the
  solver phase. For $hp$-adaptive methods, this is no longer the
  case. Consequently, balancing the cost of work between different
  processors' partitions is no longer as easy as ensuring that every
  processor owns a roughly equal number of cells. Rather, one needs to
  introduce a weighting factor for each cell that describes its
  relative cost compared to some reference. To make things worse, the
  relative cost of assembly on a cell might not match the relative
  cost of the linear solver associated with this cell, leading to
  difficult trade-offs in defining optimal weighting factors.
\end{enumerate}

All of these issues have been addressed in the current release and are
available to users. We will report on the details of the algorithms
and their performance in a separate publication.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Interfaces to the HDF5 file format and libraries}
\label{subsec:hdf5}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{GPU support via CUDA}
\label{subsec:gpu}

The GPU support was significantly extended for the current release:
\begin{itemize}
  \item \texttt{CUDAWrappers::PreconditionILU} and \texttt{CUDAWrappers::PreconditionIC}
    can be used for preconditioning \texttt{CUDAWrappers::SparseMatrix} objects.
  \item \texttt{LinearAlgebra::distributed::Vector}: the MPI-parallel vector
    class has gained a second template argument \texttt{MemorySpace} which can
    either be \texttt{Host} or \texttt{CUDA}. In the latter case, the data
    resides in the GPU memory. By default, the template parameter is
    \texttt{Host} and the behavior is unchanged compared to previous versions.
    When using CUDA, the ghost exchange can be performed either by first copying
    the relevant data to the host, performing MPI communication, and finally
    move the data to the device or, if CUDA-aware MPI is available, by
    performing MPI communication directly between GPUs.
  \item Constrained degrees of freedom: the matrix-free framework now
    supports constrained degrees of freedom. The implementation is based on the algorithms described in
    \cite{ljungkvist2017}. With this addition, both Dirichlet
    boundary conditions and the constraints arising for adaptively
    refined meshes can be imposed with in the matrix-free framework. The only restriction is that for two-dimensional meshes the
    finite element degree must be odd.  There is no such restriction in three
    dimensions.
  \item MPI matrix-free computations: using \texttt{LinearAlgebra::distributed::Vector}, the
    matrix-free framework can scale to multiple GPUs by taking
    advantage of MPI. Each MPI process can only use one GPU and therefore, if
    multiple GPUs are available in one node, it is necessary to have as many
    ranks as there are GPUs. Using Nvidia Multi-Process Service (MPS), it is also possible
    for multiple processes to use the same GPU. This can be advantageous if the
    amount of work on one rank is not sufficient to fully utilize a GPU.
\end{itemize}

The matrix-free GPU components integrated in \dealii have been compared against
CPUs in \cite{KronbichlerLjungkvist2019}, where the application to geometric
multigrid solvers is discussed.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Parallel geometric multigrid improvements}
\label{subsec:gmg}
\marginpar{performance optimizations, new features,
largest runs now at XYZ (Kronbichler?)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{New and improved tutorial and code gallery programs}
\label{subsec:steps}

talk about the new programs; also mention updating for C++11 style

\texttt{step-64} demonstrates the usage of matrix-free methods on Nvidia GPUs.
GPUs are advantageous for these kind of operations because of their superior
hardware characteristics, in particular a higher memory bandwidth than server
CPUs within a given power envelope.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Incompatible changes}

The 9.1 release includes
\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_0_1_and_9_1_0.html}
     {around 15 incompatible changes}; see \cite{changes91}. The majority of these changes
should not be visible to typical user codes; some remove previously
deprecated classes and functions, and the majority change internal
interfaces that are not usually used in external
applications. However, some are worth mentioning:
\begin{itemize}
\item The \texttt{VectorView} class was removed. We recommend either copying the
      vector subset into a \texttt{Vector} or using a \texttt{BlockVector}.
\item The function \texttt{Subscriptor::subscribe} now requires a pointer to a
      \texttt{std::atomic<bool>} that tracks whether or not the pointer to the
      subscribed-to object is still valid.
\item The \texttt{ConstraintMatrix} class gained a template parameter for scalar
      type and was been renamed \texttt{AffineConstraints}. Several methods that
      take vectors or matrices as arguments,
      such as \texttt{AffineConstraints::distribute\_local\_to\_global()},
      now require that all matrix and vector arguments have matching number
      types.
\item Similarly, the functions \texttt{create\_mass\_matrix} and
      \texttt{create\_boundary\_mass\_matrix} in the \texttt{MatrixCreator}
      namespace no longer
      support matrix and vector objects of different types.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{How to cite \dealii}\label{sec:cite}

In order to justify the work the developers of \dealii put into this
software, we ask that papers using the library reference one of the
\dealii papers. This helps us justify the effort we put into it.

There are various ways to reference \dealii. To acknowledge the use of
the current version of the library, \textbf{please reference the present
document}. For up to date information and bibtex snippets for this document
see:
\begin{center}
 \url{https://www.dealii.org/publications.html}
\end{center}

The original \texttt{\dealii} paper containing an overview of its
architecture is \cite{BangerthHartmannKanschat2007}. If you rely on
specific features of the library, please consider citing any of the
following:
\begin{itemize}
 \item For geometric multigrid: \cite{Kanschat2004,JanssenKanschat2011,ClevengerHeisterKanschatKronbichler2019};
 \item For distributed parallel computing: \cite{BangerthBursteddeHeisterKronbichler11};
 \item For $hp$ adaptivity: \cite{BangerthKayserHerold2007};
  \item For partition-of-unity (PUM) and enrichment methods of the
    finite element space: \cite{Davydov2016};
 \item For matrix-free and fast assembly techniques:
   \cite{KronbichlerKormann2012,KronbichlerKormann2019};
 \item For computations on lower-dimensional manifolds:
   \cite{DeSimoneHeltaiManigrasso2009};
 \item For integration with CAD files and tools:
   \cite{HeltaiMola2015};
 \item For Boundary Elements Computations:
   \cite{GiulianiMolaHeltai-2018-a};
 \item For \texttt{LinearOperator} and \texttt{PackagedOperation} facilities:
   \cite{MaierBardelloniHeltai-2016-a,MaierBardelloniHeltai-2016-b}.
 \item For uses of the \texttt{WorkStream} interface:
   \cite{TKB16};
   \item Fur uses of the \texttt{ParameterAcceptor} concept,
     \texttt{MeshWorker::ScratchData}, and
     \texttt{ParsedConvergenceTable}: \cite{SartoriGiulianiBardelloni-2018-a}.
\end{itemize}

\marginpar{I'm sure we have new interfaces. HDF5 and the various AD/SD
  libraries come to mind. Please update.}
\dealii can interface with many other libraries:
\begin{multicols}{3}
\begin{itemize}
\item ADOL-C \cite{Griewank1996a,adol-c}
\item ARPACK \cite{arpack}
\item Assimp \cite{assimp}
\item BLAS and LAPACK \cite{lapack}
\item cuSOLVER \cite{cusolver}
\item cuSPARSE \cite{cusparse}
\item Gmsh \cite{geuzaine2009gmsh}
\item GSL \cite{gsl2016}
\item HDF5 \cite{hdf5}
\item METIS \cite{karypis1998fast}
\item MUMPS \cite{ADE00,MUMPS:1,MUMPS:2,mumps-web-page}
\item muparser \cite{muparser-web-page}
\item nanoflann \cite{nanoflann}
\item NetCDF \cite{rew1990netcdf}
\item OpenCASCADE \cite{opencascade-web-page}
\item p4est \cite{p4est}
\item PETSc \cite{petsc-user-ref,petsc-web-page}
\item ROL \cite{ridzal2014rapid}
\item ScaLAPACK \cite{slug}
\item SLEPc \cite{Hernandez:2005:SSF}
\item SUNDIALS \cite{sundials}
\item TBB \cite{Rei07}
\item Trilinos \cite{trilinos,trilinos-web-page}
\item UMFPACK \cite{umfpack}
\end{itemize}
\end{multicols}
Please consider citing the appropriate references if you use interfaces to these
libraries.

The two previous releases of \dealii can be cited as
\cite{dealII85,dealII90}.

\section{Acknowledgments}

\dealii is a world-wide project with dozens of contributors around the
globe. Other than the authors of this paper, the following people contributed code to
this release:\\
%
% get this from the changes/*/* files using the command listed in the
% release-tasks paper and remove the authors of this paper
%
% updated 5/12/2019
\marginpar{Remove all other authors of this paper}
\marginpar{Do we want to spell out who the Gingko devs are?}
  Giovanni Alzetta,
  Mathias Anselmann,
  Daniel Appel,
  Alexander Blank,
  Vishal Boddu,
  Benjamin Brands,
  Pi-Yueh Chuang,
  T. Conrad Clevenger,
  Sambit Das,
  the Ginkgo developers,
  Stefano Dominici,
  Nivesh Dommaraju,
  Marc Fehling,
  Niklas Fehn,
  Isuru Fernando,
  Andreas Fink,
  Daniel Garcia-Sanchez,
  Rene Gassm{\"o}ller,
  Alexander Grayver,
  Joshua Hanophy,
  Logan Harbour,
  Graham Harper,
  Daniel Jodlbauer,
  Stefan Kaessmair,
  Eldar Khattatov,
  Alexander Knieps,
  Uwe K{\"o}cher,
  Kurt Kremitzki,
  Dustin Kumor,
  Ross Kynch,
  Damien Lebrun-Grandie,
  Jonathan Matthews,
  Stefan Meggendorfer,
  Pratik Nayak,
  Lei Qiao,
  Ce Qin,
  Reza Rastak,
  Roland Richter,
  Alberto Sartori,
  Svenja Schoeder,
  Sebastian Stark,
  Antoni Vidal,
  Jiaxin Wang,
  Yuxiang Wang,
  Zhuoran Wang.

Their contributions are much appreciated!

\bigskip

\dealii and its developers are financially supported through a
variety of funding sources:

D.~Arndt and M.~Kronbichler were partially supported by the German
Research Foundation (DFG) under the project ``High-order discontinuous
Galerkin for the exa-scale'' (\mbox{ExaDG}) within the priority program ``Software
for Exascale Computing'' (SPPEXA).

W.~Bangerth and T.~Heister were partially
supported by the National Science Foundation under award OAC-1835673
as part of the Cyberinfrastructure for Sustained Scientific Innovation (CSSI)
program; by award DMS-1821210; and by the Computational Infrastructure
in Geodynamics initiative (CIG), through the National Science
Foundation under Awards No.~EAR-0949446 and EAR-1550901 and The
University of California -- Davis.

D.~Davydov was supported by the German Research Foundation (DFG), grant DA
1664/2-1 and the Bayerisches Kompetenznetzwerk
f\"ur Technisch-Wissenschaftliches Hoch- und H\"ochstleistungsrechnen
(KONWIHR).

T.~Heister was also partially supported by NSF Award DMS-1522191, and
by Technical Data Analysis, Inc. through US Navy SBIR N16A-T003.

M.~Maier was partially supported by ARO MURI Award No. W911NF-14-0247.

J-P.~Pelteret was supported by the European Research Council (ERC) through
the Advanced Grant 289049 MOCOPOLY.

B.~Turcksin: This material is based upon work supported by the U.S.
Department of Energy, Office of Science, under contract number
DE-AC05-00OR22725.

D.~Wells was supported by the National Science Foundation (NSF) through Grant
DMS-1344962.

The Interdisciplinary Center for Scientific Computing (IWR) at Heidelberg
University has provided hosting services for the \dealii web page.

\bibliography{paper}{}
\bibliographystyle{abbrv}

\end{document}