### 2020-05-31 9.2/paper.tex
Added a first version of section 2.4 related t the support for particle-based methods

\documentclass{ansarticle-preprint}
%\usepackage{ucs}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
%\usepackage{cite}
\usepackage{anslistings}
\usepackage{multicol}
\usepackage{pdfsync}

\usepackage{pgfplots}
\usepackage{pgfplotstable}

\usepackage{fontenc}
\usepackage{graphicx}
\usepackage{xspace}

\usepackage{siunitx}

\usepackage{floatflt}

%\renewcommand{\baselinestretch}{2.0}

\usepackage[normalem]{ulem}

\usepackage{todonotes}

\pgfplotsset{compat=1.9}
\definecolor{gnuplot@lightblue}{RGB}{87,181,232}
\definecolor{gnuplot@green}{RGB}{0,158,115}
\definecolor{gnuplot@purple}{RGB}{148,0,212}

\newcommand{\specialword}[1]{\texttt{#1}}
\newcommand{\dealii}{{\specialword{deal.II}}\xspace}
\newcommand{\pfrst}{{\specialword{p4est}}\xspace}
\newcommand{\trilinos}{{\specialword{Trilinos}}\xspace}
\newcommand{\aspect}{\specialword{Aspect}\xspace}
\newcommand{\petsc}{\specialword{PETSc}\xspace}
\newcommand{\cmake}{{\specialword{CMake}}\xspace}
\newcommand{\autoconf}{{\specialword{autoconf}}\xspace}

%
% Author list -- please add yourself in both places below (in
%                alphabetical order) if you think that your
%                contributions to the last release warrant this
%

\hypersetup{
  pdfauthor={
     Daniel Arndt,
     Wolfgang Bangerth,
     Bruno Blais,
     Thomas C. Clevenger,
%     Denis Davydov,
    Marc Fehling,
%     Daniel Garcia-Sanchez,
    Alexander V. Grayver,
%     Graham Harper,
    Timo Heister,
    Luca Heltai,
    Martin Kronbichler,
    Peter Munch,
%     Ross Maguire Kynch,
    Matthias Maier,
    Jean-Paul Pelteret,
    Reza Rastak,
    Bruno Turcksin,
    David Wells
    Zhuoran Wang,
  },
  pdftitle={The deal.II Library, Version 9.2, 2020},
}

\title{The \dealii{} Library, Version 9.2}

 \author[*1]{Daniel Arndt}
 \affil[1]{Computational Engineering and Energy Sciences Group,
   Computional Sciences and Engineering Division,
   Oak Ridge National Laboratory, 1 Bethel Valley Rd.,
   TN 37831, USA.
   \texttt{arndtd/turcksinbr@ornl.gov}}

 \author[2]{Wolfgang~Bangerth}
 \affil[2]{Department of Mathematics and Department of Geosciences, Colorado State University, Fort
   Collins, CO 80523-1874, USA.
   \texttt{bangerth@colostate.edu}}

\author[3]{Bruno Blais}
\affil[3]{Research Unit for Industrial Flows Processes (URPEI), Department of Chemical Engineering,
          Polytechnique Montréal,
          PO Box 6079, Stn Centre-Ville, Montréal, Québec, Canada, H3C 3A7.
          {\texttt{bruno.blais@polymtl.ca}}}

 \author[4]{Thomas~C.~Clevenger}
 \affil[4]{School of Mathematical and Statistical Sciences,
   Clemson University,
   Clemson, SC, 29634, USA
   {\texttt{tcleven/heister@clemson.edu}}}
%
% \author[4]{Denis~Davydov}
% \affil[4]{Chair of Applied Mechanics,
%   Friedrich-Alexander-Universit\"{a}t Erlangen-N\"{u}rnberg,
%   Egerlandstr.\ 5,
%   91058 Erlangen, Germany.
%   {\texttt{\{denis.davydov,jean-paul.pelteret\}@fau.de}}}
%
\author[5]{Marc~Fehling}
\affil[5]{Institute for Advanced Simulation,
  Forschungszentrum J\"{u}lich GmbH,
  52425 J\"{u}lich, Germany.
  {\texttt{m.fehling@fz-juelich.de}}}
%
% \author[6]{Daniel Garcia-Sanchez}
% \affil[6]{Sorbonne Universit\'es, UPMC Univ.\ Paris 06, CNRS-UMR 7588,
%   Institut des NanoSciences de Paris, F-75005, Paris, France
%   {\texttt{daniel.garcia-sanchez@insp.upmc.fr}}}
%
% \author[2]{Graham Harper}
%

\author[6]{Alexander~V.~Grayver}
\affil[6]{Institute of Geophysics,
  ETH Zurich,
  Sonneggstrasse 5, 8092 Z\"{u}rich, Switzerland.
  {\texttt{agrayver@ethz.ch}}}

\author[4]{Timo~Heister}

\author[7]{Luca~Heltai}
\affil[7]{SISSA,
   International School for Advanced Studies,
   Via Bonomea 265,
   34136, Trieste, Italy.
   {\texttt{luca.heltai@sissa.it}}}

 \author[8]{Martin~Kronbichler}
 \affil[8]{Institute for Computational Mechanics,
   Technical University of Munich,
   Boltzmannstr.~15, 85748 Garching, Germany.
   {\texttt{kronbichler/munch@lnm.mw.tum.de}}}

\author[9]{Matthias~Maier}
\affil[9]{Department of Mathematics,
  Texas A\&M University,
  3368 TAMU,
  College Station, TX 77845, USA.
  {\texttt{maier@math.tamu.edu}}}

\author[8,10]{Peter Munch}
 \affil[10]{Institute of Materials Research, Materials Mechanics,
 Helmholtz-Zentrum Geesthacht,
 Max-Planck-Str. 1, 21502 Geesthacht, Germany.
   {\texttt{peter.muench@hzg.de}}}

\author[11]{Jean-Paul~Pelteret}
\affil[11]{Independent researcher.
{\texttt{jppelteret@gmail.com}}}

 \author[12]{Reza Rastak}
 \affil[12]{Department of Civil and Environmental Engineering,
   Stanford University,
   Stanford, CA 94305, USA.
   {\texttt{rastak@stanford.edu}}}

 \author[**13]{Ignacio Tomas}
 \affil[13]{Sandia National Laboratories, Org 1442,
 P.O. Box 5800, MS 1320, Albuquerque, NM, 87185-1320,
   \texttt{itomas@sandia.gov}}

\author[*1]{Bruno~Turcksin}
%
\author[14]{David Wells}
\affil[14]{Department of Mathematics, University of North Carolina,
  Chapel Hill, NC 27516, USA.
  {\texttt{drwells@email.unc.edu}}}

\author[15]{Zhuoran Wang}
\affil[15]{Department of Mathematics, Colorado State University, Fort Collins,
          CO 80523-1874, USA.
{\texttt{zhrwang@math.colostate.edu}}}
\renewcommand{\labelitemi}{--}

\begin{document}
\maketitle

\footnotetext{%
  $^\ast$ This manuscript has been authored by UT-Battelle, LLC under Contract No.
  DE-AC05-00OR22725 with the U.S. Department of Energy. The United States
  Government retains and the publisher, by accepting the article for
  publication, acknowledges that the United States Government retains a
  non-exclusive, paid-up, irrevocable, worldwide license to publish or reproduce
  the published form of this manuscript, or allow others to do so, for United
  States Government purposes. The Department of Energy will provide public
  access to these results of federally sponsored research in accordance with the
  DOE Public Access Plan (http://energy.gov/downloads/doe-public-access-plan).}

\footnotetext{%
  $^{**}$ Sandia National Laboratories is a multimission laboratory managed and
  operated by National Technology \& Engineering Solutions of Sandia, LLC, a
  wholly owned subsidiary of Honeywell International Inc., for the U.S.
  Department of Energy's National Nuclear Security Administration under contract
  DE-NA0003525. This document describes objective technical results and analysis.
  Any subjective views or opinions that might be expressed in the paper do not
  necessarily represent the views of the U.S. Department of Energy or the United
  States Government.}

\begin{abstract}
  This paper provides an overview of the new features of the finite element
  library \dealii, version 9.2.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Overview}

\dealii{} version 9.2.0 was released May XYZ, 2020.
\todo[inline]{Need to update date.}
This paper provides an
overview of the new features of this release and serves as a citable
reference for the \dealii{} software library version 9.2. \dealii{} is an
object-oriented finite element library used around the world in the
development of finite element solvers. It is available for free under the
GNU Lesser General Public License (LGPL). Downloads are available at
\url{https://www.dealii.org/} and \url{https://github.com/dealii/dealii}.

The major changes of this release are:
%
\begin{itemize}
\item xy \todo[inline]{Update once we have the subsections in Section 2;
  provide cross-references to each of these subsections}
%   \item Improved support for automatic differentiation (see
%     Section~\ref{subsec:ad}),
\end{itemize}
%
These major changes are discussed in detail in Section~\ref{sec:major}. There
are a number of other noteworthy changes in the current \dealii{} release
that we briefly outline in the remainder of this section:
%
\begin{itemize}
\item \dealii{} had decent support for solving complex-valued problems
  for a while already
  (e.g., ones in quantum mechanics -- like the equation used in the
  \texttt{step-58} tutorial program covered below -- or for
  time-harmonic problems). However, there were two
  areas in which support was missing. First, the UMFPACK direct solver
  packaged with \dealii{} did not support solving complex-valued
  linear problems. This has now been addressed: UMFPACK actually can
  solve such systems, we just needed to write the appropriate
  interfaces. Second, the \texttt{DataOut} class that is responsible
  for converting nodal data into information that can then be written
  into files for visualization did not know how to deal with vector-
  and tensor-valued fields whose components are complex numbers. An
  example for this is to solve the time-harmonic version of the
  Maxwell equations that has the electric and magnetic fields as
  solution. This, too, has been addressed in this release.
\item The new \texttt{DiscreteTime} class provides a more
  consistent, more readable, and less error-prone approach to control
  time-stepping algorithms within time-dependent simulations.
  The interface of this class is designed to be minimal
  to enforce a number of important programming invariants, reducing
  the possibility of mistakes in the user code.
  An example is that \texttt{DiscreteTime} ensures that the final time step ends
  precisely on a predefined end time, automatically
  lengthening or shortening to final time step.
\item A key component of \dealii{} are the \texttt{FEValues} and
  \texttt{FEFaceValues} classes that evaluate finite element functions
  at quadrature points located on cells and faces of a cell,
  respectively~\cite{BangerthHartmannKanschat2007}. This release now
  contains a class \texttt{FEInterfaceValues} that considers the
  restriction of the shape functions from \textit{both} sides of a face and
  allows evaluating jumps and averages of shape functions along this
  face. These are common components of the bilinear forms of
  discontinuous Galerkin schemes (as well as schemes for fourth-order
  equations, see the discussion of \texttt{step-47} below) and greatly
  simplify the implementation of these methods.
\end{itemize}
%
The changelog lists more than 240 other
features and bugfixes.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Major changes to the library}
\label{sec:major}

This release of \dealii{} contains a number of large and significant changes
that will be discussed in this section.
It of course also contains a
vast number of smaller changes and added functionality; the details of these
can be found
\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_1_1_and_9_2_0.html}{
in the file that lists all changes for this release}, see \cite{changes92}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{A new fully distributed triangulation class}
\label{subsec:pft}

Previously, all triangulation classes of \dealii{} had in common that the coarse grid is replicated by
all processes in a parallel environment, and the actual mesh used for computations is constructed by repeated
refinement. However, this has its limitations in many
applications where the mesh comes
from an external mesh generator in the form of a file that frequently
already contains millions
or tens of millions of cells. For such configurations, applications might exhaust
available memory already while reading the mesh on each MPI process.

The new \texttt{parallel::fullydistributed::Triangulation} class targets this issue
by distributing also the coarse grid. Such
a triangulation can be created by providing to each process a \texttt{Triangulation\-De\-scrip\-tion::Description} struct, containing
(i) the relevant data to construct the local part of the coarse grid, (ii) the
translation of the local coarse-cell IDs to globally unique IDs, (iii) the hierarchy
of mesh refinement steps, and (iv) the owner of the cells on the active mesh level as well
as on the multigrid levels. For the current release, triangulations set up
this way cannot be adaptively refined after construction, though we plan to
improve this for the next release.

%The \texttt{TriangulationDescription::Description} struct can be filled manually or
%by the utility functions from the \texttt{TriangulationDescription::Utilities}
%namespace. The function \texttt{create\_\allowbreak description\_\allowbreak
%from\_ \allowbreak triangulation()} can convert a base triangulation (partitioned
%serial \texttt{Tri\-angulation} and \texttt{parallel::distributed::Triangulation})
%to such a struct. The advantage of this approach is that all known utility
%functions from the namespaces \texttt{GridIn} and \texttt{GridTools} can be used
%on the base triangulations before converting them to the structs. Since this
%function suffers from the same main memory problems as described above, we also
%provide the function \texttt{create\_description\_from\_triangulation\_in\_groups()},
%which creates the structs only on the master process in a process group. These
%structs are filled one by one and are sent to the relevant processes once they are
%ready. A sensible process group size might contain all processes of one compute node.

The new fully distributed triangulation class supports 1D, 2D, and 3D meshes
including geometric multigrid hierarchies, periodic boundary conditions, and
hanging nodes.

%We intend to extend the usability of the new triangulation class in regard of
%different aspects, e.g., I/O. In addition, we would like to enable adaptive mesh
%refinement, a feature of the other triangulation classes, which is very much
%appreciated by many users. For repartitioning, we plan to use an oracle approach
%known from \texttt{parallel::distributed::Triangulation}. Here, we would like to
%rely on a user-provided partitioner, which might also be a graph partitioner.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Improved large-scale performance}
\label{subsec:performance}

Large-scale simulations with up to 304,128 cores have revealed bottlenecks in release
9.1 during initialization of a number of distributed data structures, due to the usage of expensive collective operations
like \texttt{MPI\_Allgather()} and \texttt{MPI\_\allowbreak
  Alltoall()}. Typical examples are the
pre-computation of the indices of those vector entries (or
other linear index ranges) owned by
each process, which were previously stored in  an array on every process.
This information is needed to set up
the  \texttt{Utilities::MPI::Par\-ti\-ti\-oner} class.
In release 9.2, we have replaced these functions in favor of
consensus algorithms~\cite{hoefler2010scalable}, which can be
found in the namespace \texttt{Utilities::\allowbreak MPI::\allowbreak ConsensusAlgorithms} (short: \texttt{CA}).
Now, only the locally relevant information about the index ranges is
(re)computed when needed, which, for more than 100 MPI processes, uses
point-to-point communications and a single \texttt{MPI\_IBarrier()}.

%Consensus algorithms are algorithms dedicated to efficient dynamic-sparse
%communication patterns. In this context, the term ``dynamic-sparse'' means
%that by the time this function is called, the other processes do not know
%yet that they have to answer requests and
%each process only has to communicate with a small subset of processes of the
%MPI communicator. We provide two flavors of the consensus algorithm: the two-step
%approach \texttt{ConsensusAlgorithms::PEX} and the \texttt{ConsensusAlgorithms::NBX},
%which uses only point-to-point communications and a single \texttt{MPI\_IBarrier()}.
%The class \texttt{ConsensusAlgorithms::Selector} selects one of the two previous
%algorithms, depending on the number of processes.

Users can apply the new algorithms for their own dynamic-sparse problems by
providing a list of target
processes and pack/unpack routines either by implementing the interface
\texttt{CA::\allowbreak Process} or by providing \texttt{std::function}
objects to \texttt{CA::AnonymousProcess}.

\todo[inline]{Aren't the target processes the result of the operation, and the
  input global numbers of indices that we request?}

By replacing the collective communications during set up and removing the arrays
that contain information for each process (enabled by the application of consensus
algorithms and other modifications---a full list of modifications leading to this
improvement can be found online), we were able to significantly
improve the set up
time for large-scale simulations and to solve a Poisson problem with multigrid
with \num{2.1e12} unknowns.
Figure~\ref{fig:init_costs} compares timings of simulations of various problem
sizes (including set up) on 49,152 MPI ranks using a matrix-free
solver~\cite{KronbichlerKormann2019,KronbichlerWall2018}; this solver uses discontinuous elements of
degree $5$ in a geometric multigrid (GMG) scheme. The comparison between the previous release 9.1
and the current release 9.2 shows that while the scaling for the V-cycle had been
very good before, many initialization routines have been considerably
improved, especially the enumeration of unknowns on the multigrid levels and
the setup of the multigrid transfer.

\pgfplotstableread{
dofs  grid  distdofs  mfsetup  mgtransfer  smoother  vcycle
14155776  0.347457  0.02424084  0.03346427  0.0105426  0.0108817  0.0035033
28311552  0.35134  0.0215024  0.0340963  0.0095201  0.010616  0.0035887
56623104  0.41577  0.0243593  0.0432297  0.011557  0.011667  0.0041244
113246208  0.42172  0.0323919  0.0427806  0.012983  0.011834  0.0043853
226492416  0.43565  0.0404079  0.0421606  0.012867  0.017305  0.0056340
452984832  0.5055  0.0500101  0.0562977  0.017739  0.022763  0.0061992
905969664  0.54552  0.065865  0.0641994  0.019445  0.037431  0.0087932
1811939328  0.5672  0.104909  0.080807  0.025395  0.063548  0.0133252
3623878656  0.65541  0.165071  0.129717  0.042294  0.12644  0.0218619
7247757312  0.74521  0.251417  0.221072  0.073314  0.21807  0.0339264
14495514624  0.84789  0.403463  0.36894  0.11425  0.4091  0.0669798
28991029248  0.91983  0.52435  0.5363  0.14189  0.79178  0.1322448
57982058496  1.0339  0.87448  0.93134  0.23863  1.5896  0.2539445
115964116992  1.2667  1.46016  1.70806  0.44101  3.1409  0.4956625
231928233984  1.7597  2.58121  3.3741  0.77624  6.2628  0.9857200
}\tableinit

\pgfplotstableread{
dofs  grid  distdofs  mfsetup  mgtransfer  smoother  vcycle
14155776  2.0148  0.9086  0.433886  1.3099  0.010685  0.0035033
28311552  1.8746  0.91078  0.576777  1.3093  0.010752  0.0035887
56623104  2.0924  0.98058  0.574183  1.4459  0.011831  0.0041244
113246208  2.3788  1.22025  0.547492  1.7919  0.011979  0.0043853
226492416  2.1254  1.50329  0.839219  1.7832  0.017819  0.0056340
452984832  2.4325  2.10171  0.711708  2.0263  0.023393  0.0061992
905969664  3.8911  3.58355  1.08478  2.5844  0.045697  0.0087932
1811939328  4.1803  4.379  0.784804  1.8947  0.061552  0.0133252
3623878656  2.6654  5.6073  0.833896  1.8912  0.1188  0.0218619
7247757312  2.7422  7.8463  0.741733  1.8477  0.21291  0.0339264
14495514624  2.765  9.9848  0.99218  1.8248  0.40665  0.0669798
28991029248  3.2901  15.0526  1.5124  2.6723  0.81013  0.1322448
57982058496  3.1927  21.0562  1.72831  2.1998  1.5905  0.2539445
115964116992  4.1485  31.0087  2.67998  2.9091  3.2009  0.4956625
231928233984  4.1787  47.195  4.2821  3.9558  6.3362  0.9857200
}\tableinitold

\begin{figure}
  \begin{tikzpicture}
    \begin{loglogaxis}[
      width=0.51\textwidth,
      height=0.38\textwidth,
          xlabel={\# DoFs},
          ylabel={time [s]},
          title={9.1 release},
          xtick={1.4e7,5.7e7,2.3e8,9.1e8,3.6e9,1.4e10,5.8e10,2.3e11},
          xticklabels={1.4$\times 10^7$,5.7$\times 10^7$,2.3$\times 10^8$,9.1$\times 10^8$,3.6$\times 10^9$,1.4$\times 10^{10}$,5.8$\times 10^{10}$,2.3$\times 10^{11}$},
          tick label style={font=\scriptsize},
          xticklabel style={rotate=25,anchor=north east},
          label style={font=\scriptsize},
          legend style={font=\scriptsize},
          %legend to name=legend:scalingMG,
          %legend cell align=left,
          %legend columns = 3,
          ymin=2e-3,ymax=8e1,
          grid, thick]
          \addplot table[x index = 0, y index={1}] {\tableinitold};
          \addplot[red,mark=square*] table[x index = 0, y index={2}] {\tableinitold};
          \addplot[gnuplot@green,mark=otimes] table[x index = 0, y index={3}] {\tableinitold};
          \addplot[gnuplot@lightblue,mark=triangle] table[x index = 0, y index={4}] {\tableinitold};
          \addplot[gnuplot@purple,mark=diamond] table[x index = 0, y index={5}] {\tableinitold};
          \addplot[black,mark=x] table[x index = 0, y index={6}] {\tableinitold};
    \end{loglogaxis}
  \end{tikzpicture}
  \hfill
  \begin{tikzpicture}
    \begin{loglogaxis}[
      width=0.51\textwidth,
      height=0.38\textwidth,
          xlabel={\# DoFs},
          ylabel={time [s]},
          title={9.2 release},
          xtick={1.4e7,5.7e7,2.3e8,9.1e8,3.6e9,1.4e10,5.8e10,2.3e11},
          xticklabels={1.4$\times 10^7$,5.7$\times 10^7$,2.3$\times 10^8$,9.1$\times 10^8$,3.6$\times 10^9$,1.4$\times 10^{10}$,5.8$\times 10^{10}$,2.3$\times 10^{11}$},
          tick label style={font=\scriptsize},
          xticklabel style={rotate=25,anchor=north east}, thick,
          label style={font=\scriptsize},
          legend style={font=\scriptsize},
          legend to name=legend:init,
          legend cell align=left,
          legend columns = 6,
          ymin=2e-3,ymax=8e1,
          grid]
          \addplot table[x index = 0, y index={1}] {\tableinit};
          \addlegendentry{create mesh};
          \addplot[red,mark=square*] table[x index = 0, y index={2}] {\tableinit};
          \addlegendentry{enumerate (mg) dofs};
          \addplot[gnuplot@green,mark=otimes] table[x index = 0, y index={3}] {\tableinit};
          \addlegendentry{init matrix-free};
          \addplot[gnuplot@lightblue,mark=triangle] table[x index = 0, y index={4}] {\tableinit};
          \addlegendentry{init GMG transfer};
          \addplot[gnuplot@purple,mark=diamond] table[x index = 0, y index={5}] {\tableinit};
          \addlegendentry{init smoother};
          \addplot[black,mark=x] table[x index = 0, y index={6}] {\tableinit};
          \addlegendentry{GMG V-cycle};
    \end{loglogaxis}
  \end{tikzpicture}
  \\
  \strut\hfill\pgfplotslegendfromname{legend:init}\hfill\strut
  \caption{\it Comparison of initialization costs of various data structures in the 9.1 release (left) and the new 9.2 release (right) when run on 49,152 MPI ranks.}
  \label{fig:init_costs}
\end{figure}

\todo[inline]{Timo: Do you want to say something about that the new code has been also applied to solve problems with adaptively refined
meshes with more than \num{4e9} unknowns?}

\todo[inline]{Timo says: We did solve adaptive problems with more than
  4B unknowns before, but some were broken for a while during this
  release stage due to the consensus algorithms. But we did not test
  systems or block vectors before. But we should not make it sound too
  negative.}

\todo[inline]{Wolfgang says: Should we just not mention any of this at
  all? The section stands well on its own.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Better support for parallel $hp$-adaptive algorithms}
\label{subsec:hp}

Since the previous release, \dealii{} has had support for $hp$-adaptive finite element methods
on distributed memory systems \cite{dealII91}. We implemented the bare functionality
for $hp$-adaptive methods with the objective to offer the greatest flexibility in
their application. Here, reference finite elements still had to be assigned manually
to each cell, which may not lead to an optimal mesh and is tedious.

With the current release, $hp$-adaptive finite element methods have been further
expanded: New features like decision strategies have been added and the user interface
has been overhauled, effectively making $hp$-methods more attractive to use. We introduced
many new functions that automatize the general workflow for applying $hp$-decision
strategies, which run on top of the previous low-level implementation for both serial
and parallel applications.

The interface is now as simple to use as the one for $h$-adaptive mesh refinement.
Consider the following (incomplete) listing as an example: We estimate both error and
smoothness of the finite element approximation. Further, we flag certain fractions of
cells with the highest and lowest errors for refinement and coarsening, respectively
(here: 30\%/3\%). From those cells listed for adaptation, we designate a subset
for $h$- and $p$-adaptation (here: 10\%/90\%). The parameters of the corresponding
\texttt{hp::Refinement} function specify the fraction of cells to be $p$-adapted from
those subsets flagged for refinement and coarsening, respectively.
\begin{c++}
Vector<float> estimated_error_per_cell(n_active_cells);
KellyErrorEstimator::estimate(
  dof_handler, ..., solution, estimated_error_per_cell, ...);
GridRefinement::refine_and_coarsen_fixed_number(
  triangulation, estimated_error_per_cell, 0.3, 0.03);

Vector<float> estimated_smoothness_per_cell(n_active_cells);
SmoothnessEstimator::Legendre::coefficient_decay(
  ..., dof_handler, solution, estimated_smoothness_per_cell);
hp::Refinement::p_adaptivity_fixed_number(
  dof_handler, estimated_smoothness_per_cell, 0.9, 0.9);
hp::Refinement::choose_p_over_h(dof_handler);

triangulation.execute_coarsening_and_refinement();
\end{c++}

In particular, we implemented decision strategies based on refinement history and
smoothness estimation, and made sure that they work for refinement as well as
coarsening in terms of $h$- and $p$-adaptation in serial and parallel applications.

The former relies on knowing an estimate for the upper error bound \cite[Thm.~3.4]{BabuskaSuri1990}.
For successive refinements, we can predict how the error will change based on
current error estimates and adaptation flags. In the next refinement cycle, these
predicted error estimates allow us to decide whether the choice of adaptation in
the previous cycle was justified, and provide a criterion for the choice in the next
cycle \cite{MelenkWohlmuth2001}.

In general, $p$-refinement is favorable over $h$-refinement in smooth regions of
the finite element approximation \cite[Thm.~3.4]{BabuskaSuri1990}. Thus, estimating
its smoothness provides a suitable decision indicator for $hp$-adaptation. For this
purpose, we express the finite element approximation in an orthogonal
basis of increasing frequency, and consider the decay of their expansion
coefficients as the estimation of smoothness. This has been implemented for both
Fourier coefficients \cite{BangerthKayserHerold2007} and Legendre coefficients
\cite{Mavriplis1994,HoustonSeniorSueli2003,HoustonSueli2005,EibnerMelenk2007}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Support for particle-based methods}
\label{subsec:particles}

Support for particles was originally introduced in \dealii{} version 9.0. In \dealii,
particles can be used as passive tracers, or as part of more complex
models such as those based on Particle-In-Cells (PIC) approaches \cite{gassmoller2018flexible}.

With the current release, the support for particles has been further expanded: New parallel
 insertion mechanisms and a basic interface to post-process particles have
been added, effectively making their usage more flexible and enabling a larger
range of use cases (e.g. immersed boundaries in step-70).

Through the addition of the \texttt{Particles::ParticleHandler::insert\_global\_particles} member function,
 particles can now be inserted in parallel from a vector of points even if these points do not lie
 on the subdomain from which the insertion is called. This operation requires extensive 
 communication between the processes to locate   the MPI process the cell in which the particle
  is located. However, it is made significantly faster  through the usage of bounding boxes
  that surround the cells of each subdomain. 
  This function also takes care of transferring the properties attached to the particles to their new owner.
 This new capacity enables particle generators that insert particles at the location
 of the support points (\texttt{Particles::Generator::dof\_support\_points}) and at the quadrature points (\texttt{Particles::Generator::quadrature\_points}) of a
 possibly non-matching triangulation. Consequently, complex particle patterns can be inserted
 using unstructured hexahedral grids generated outside of \dealii{}.

To visualize the motion of particles, the \texttt{Particles::DataOut} class was added to the library.
In its current version, this class supports writing the particles positions and id in all format supported
by \dealii, but does not allow writing the properties attached to the particles.

\todo[inline]{Luca: Your section}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Improved performance of the symbolic differentiation framework}
\label{subsec:symbdiff}

\todo[inline]{J-P's section}

BatchOptimizer for symbolic expressions
\begin{itemize}
\item CSE for dictionary-based expressions
\item ``lambdify'' (via SymEngine)
\item Offloading to LLVM JIT compiler (via SymEngine)
\item Serialization, so complex expressions can be compiled offline
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Advances of the SIMD capabilities and the matrix-free infrastructure}
\label{subsec:mf}

%\begin{itemize}
%\item ECL
%\item VectorizedArrayType
%\end{itemize}

The class \texttt{VectorizedArray<Number>} is a key component to achieve
the high node-level performance of the matrix-free algorithms in
\dealii{}~\cite{KronbichlerKormann2012, KronbichlerKormann2019}. It is a
wrapper class around a short vector of $n$ entries of type \texttt{Number}
and maps arithmetic operations to appropriate
single-instruction/multiple-data (SIMD) concepts by intrinsic functions.
The class \texttt{VectorizedArray} has been made more user-friendly in this
release by making it compatible with the STL algorithms found in the header
\texttt{<algorithm>}. The length of the  vector can now be queried by
\texttt{VectorizedArray::size()} and its underlying number type by
\texttt{VectorizedArray::value\_type}. The \texttt{VectorizedArray} class
now supports range-based iteration over its entries. In addition \dealii{}
now supports ternary operations on vectorized data, where for a given
binary comparison operation a true or false value is selected. For example,
the vectorized equivalent of \texttt{(left < right) ? true\_value :
false\_value} can be expressed as
\begin{c++}
auto result = compare_and_apply_mask<SIMDComparison::less_than>(
  left, right, true_value, false_value);
\end{c++}
which compares every element of the vectorized array individually and
selects the corresponding value from the \texttt{true\_value}, or
\texttt{false\_value} array.

In previous \dealii{} releases, the vector length was set at compile time
of the library to match the highest value supported by the given processor
architecture. Now, a second optional template argument can be specified as
\texttt{VectorizedArray<Number, size>}, where \texttt{size} explicitly
controls the vector length within the capabilities of a particular
instruction set. (A full list of supported vector lengths is presented in
Table~\ref{tab:vectorizedarray}.) This allows users to select the vector
length/ISA and, as a consequence, the number of cells to be processed at
once in matrix-free operator evaluations. For example, the \dealii{}-based
library \texttt{hyper.deal}~\cite{munch2020hyperdeal}, which solves the 6D
Vlasov--Poisson equation with high-order discontinuous Galerkin methods
(with more than a thousand degrees of freedom per cell), constructs a
tensor product of two \texttt{MatrixFree} objects of different SIMD-vector
length in the same application and benefits---in terms of performance---by
the possibility of decreasing the number of cells processed by a single
SIMD instruction.

\begin{table}
\caption{\it Supported vector lengths of the class \texttt{VectorizedArray} and
the corresponding instruction-set-architecture extensions. }\label{tab:vectorizedarray}
\centering
\begin{tabular}{ccc}
\toprule
\textbf{double} & \textbf{float} & \textbf{ISA}\\
\midrule
VectorizedArray<double, 1> & VectorizedArray<float, 1> & (auto-vectorization) \\
VectorizedArray<double, 2> & VectorizedArray<float, 4> & SSE2/AltiVec \\
VectorizedArray<double, 4> & VectorizedArray<float, 8> & AVX/AVX2 \\
VectorizedArray<double, 8> & VectorizedArray<float, 16> & AVX-512 \\
\bottomrule
\end{tabular}

\caption{\it Comparison of relevant SIMD-related classes in \dealii{} and \texttt{C++23}.}\label{tab:simd}
\centering
\begin{tabular}{cc}
\toprule
\textbf{VectorizedArray (\dealii{})} & \textbf{std::simd (\texttt{C++23})} \\
\midrule
VectorizedArray<Number> & std::experimental::native\_simd<Number> \\
VectorizedArray<Number, size> & std::experimental::fixed\_size\_simd<Number, size> \\ \bottomrule
\end{tabular}
\end{table}

The new interface of \texttt{VectorizedArray} also enables replacement
by any type with a matching interface. Specifically, this prepares
\dealii{} for the \texttt{std::simd} class that is slated to become
part of the \texttt{C++23} standard.
Table~\ref{tab:simd} compares the \dealii{}-specific SIMD classes and
the equivalent \texttt{C++23} classes. These changes also prepare for specialized
code paths exploiting
vectorization within an element, see~\cite{KronbichlerKormann2019}.

%We welcome the standardization of the SIMD
%parallelization paradigm in C++ and intend to replace step by step our own
%wrapper class, which has been continuously developed over the last decade. We
%would like to emphasize that the work invested in this class was not in vain,
%since many performance-relevant utility functions implemented with \texttt{VectorizedArray} in mind (e.g., \texttt{vectorized\_load\_and\_transpose}
%and \texttt{vectorized\_transpose\_and\_store}) will be still used, since they
%have not become part of the standard.

%Further additions to the \texttt{MatrixFree} infrastructure consist of:
%\begin{itemize}
%\item a new variant of \texttt{MatrixFree::cell\_loop()}: It takes two
%\texttt{std::function} objects with ranges on the locally owned degrees of freedom, one
%with work to be scheduled before the cell operation first touches some
%unknowns and another with work to be executed after the cell operation last
%touches them. The goal of
%these functions is to bring vector operations close to the time when the
%vector entries are accessed by the cell operation, which increases the cache
%hit rate of modern processors by improved temporal locality.
%\item a new form of loop \texttt{MatrixFree::loop\_cell\_centric()}: This
%kind of loop can be used in the context of discontinuous Galerkin methods,
%where both cell and face integrals have to be evaluated. While in the case of
%the traditional \texttt{loop}, cell and face integrals have been performed
%independently, the new loop performs all cell and face integrals of a cell in
%one go. This includes that each face integral has to be evaluated twice, but
%entries have to be written into the solution vector only once with improved
%data locality. Previous publications based on \dealii{} have shown the
%relevance of the latter aspect for reaching higher performance.
%\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Advances in GPU support}
\label{subsec:gpu}

For this release, the most noteworthy improvements in GPU support are the
simplification of the kernel written by user, improvement of error messaging,
and overlapping of computation and communication when using CUDA-aware MPI with
the matrix-free framework.

In order to simplify user code, we now recompute
local degree of freedom and quadrature point indices instead of having the user
keeping track of them. A new option to overlap computation and communication is
now available when using CUDA-aware MPI and matrix-free. The underlying idea is
that only the degrees of freedom (DoFs) on the boundary of the local domain
require communication. Before evaluating the matrix-free operator at these
degrees of freedom, we need to communicate ghost DoFs from other
processors.  Similarly once the operator has been evaluated, we need to update
the resulting global vector with values from other processors. The strategy that
we are now using consists in splitting the DoFs into three groups: one group of DoFs
that are on the local boundary and two groups each owning half of the interior
DoFs. When evaluating the matrix-free operator, we start the MPI communication
to get the ghosted DoFs and evaluate the operator on one of the two interior
DoFs group. When the evaluation is done, we wait for the MPI
communication to be over, and
then evaluate the operator on the boundary DoFs group. When this evaluation is
completed, we start the communication to update the global vector, we evaluate the
operator on the last group of DoFs, and finally wait for the MPI
communication to finish.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Expanded use of C++11 facilities}
\label{subsec:cxx}

Certain types of quantities in a simulation are constants fully known
at compile time. They can be pre-calculated and stored in the
compiled executable in order to avoid unnecessary initialization during
runtime. C++11 and later standards enables such computations by
marking variables and functions with the \texttt{constexpr} keyword.

This optimization is now enabled for the class templates
\texttt{Tensor} and \texttt{SymmetricTensor}.
For instance, the linear mechanical constitutive model for isotropic elastic solids
uses a constant fourth-order elasticity tensor
$\mathbb{C} = \lambda \boldsymbol{I} \otimes \boldsymbol{I} + 2 \mu \mathbb{I}$
which does not depend on the current state of strain.
This tensor can now be statically initialized by defining it as
\texttt{constexpr SymmetricTensor<4, dim>}.
As another example, the lattice vectors in a crystal plasticity model are generally
constant and known during compilation time, enabling their efficient definition as
\texttt{constexpr Tensor<1, dim>}.

Declaring variables, functions, and methods as \texttt{constexpr}
is a C++11 feature that was later expanded by the C++14 standard.
Thus, parts of the \texttt{constexpr} support in \dealii{} depend on the C++
standard supported by the compiler used to install the library.

The next release of \dealii{} will require compiler support for the C++14 standard.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{New and improved tutorial and code gallery programs}
\label{subsec:steps}

Many of the \dealii{} tutorial programs were revised in a variety of
ways as part of this release. A particular example is that we have
converted a number of programs to use range-based for loops (a C++11
feature) for loops over a range of integer indices such as loops over
all quadrature points or all indices of degrees of freedom during
assembly. This makes sense given that the
range-based way of writing loops seems to be the idiomatic approach
these days, and that we had previously already converted loops over
all cells in this way.

In addition, there are a number of new tutorial programs:
\begin{itemize}
\item \texttt{step-47} is a new program that solves the biharmonic
  equation $\Delta^2 u = f$ with
  ``clamped'' boundary condition given by $u=g, \partial u/\partial
  n=h$. This program is based on the $C^0$ interior
  penalty ($C^0$IP) method for fourth order problems
  \cite{Brenner2005}. In order to overcome
  shortcomings of classical approaches, this method uses $C^0$ Lagrange finite
  elements and introduces ``jump'' and ``average'' operators on
  interfaces of elements that penalize the jump of the gradient of the
  solution in order to obtain convergence to the $H^2$-regular
  solution of the equation.

  The $C^0$IP approach is a modern alternative to classical methods that
  use $C^1$-conforming elements such as the Argyris
  element, the Clough-Tocher element and others, all developed in the
  late 1960s. From a twenty-first century perspective, they can only be
  described as bizarre in their construction. They are also exceedingly
  cumbersome to implement if one wants to use general meshes. As a
  consequence, they have largely fallen out of favor and \dealii{} currently
  does not contain implementations of these shape functions.

\item \texttt{step-50}
  \todo[inline]{Timo/Conrad/... to write}

\item \texttt{step-58} is a program that solves the nonlinear
  Schr{\"o}dinger equation, which in non-dimensional form reads
  \begin{align*}
  - i \frac{\partial \psi}{\partial t}
  - \frac 12 \Delta \psi
  + V \psi
  + \kappa |\psi|^2 \psi
  &= 0,
  \end{align*}
  augmented by appropriate initial and boundary conditions and using
  an appropriate form for the potential $V=V(\mathbf x)$. The
  tutorial program focuses on two specific aspects for which this
  equation serves as an excellent test case: (i) Solving
  complex-valued problems without splitting the equation into its
  real and imaginary parts (as \texttt{step-29} does, for
  example). (ii)~Using operator splitting techniques. The equation is
  a particularly good test case for this technique because the only
  nonlinear term, $\kappa |\psi|^2 \psi$, does not contain any
  derivatives and consequently forms an ODE at each node point, to be
  solved in each time 
  step in an operator splitting scheme (for which, furthermore, there
  exists an analytic solution), whereas the remainder of the
  equation is linear and easily solved using standard finite element
  techniques.

\item \texttt{step-65} presents \texttt{TransfiniteInterpolationManifold}, a
manifold class that can propagate curved boundary information into the
interior of a computational domain by transfinite interpolation \cite{Gordon82}.
This manifold is a prototype for many other manifolds in that it is relatively
expensive to compute the new points, especially for higher order mappings. Since
typical programs query higher-order geometries in a large variety of contexts,
the contribution of the mapping to the run time can be significant. As a
solution, the tutorial also presents the class \texttt{MappingQCache}, which
samples the information of expensive manifolds in the points of a
\texttt{MappingQ} and caches it. The tutorial shows that this makes all queries
to the geometry very cheap.

\item \texttt{step-67} is an explicit time integrator for the
compressible Euler equations discretized with a high-order discontinuous
Galerkin (DG) scheme using the matrix-free infrastructure. Besides the use of
matrix-free evaluators for systems of equations and over-integration, it also
presents \texttt{MatrixFreeOperators::CellwiseInverseMassMatrix}, a fast implementation
of the action of the inverse mass matrix in the DG setting using tensor
products. Furthermore, this tutorial demonstrates the usage of new
pre and post operations, which can be passed to
\texttt{MatrixFree::cell\_loop()}, to schedule operations on sections of vectors close
to the matrix-vector product to increase data locality.

\item \texttt{step-69} presents a first-order scheme solving compressible
Euler's equations of gas dynamics with a graph-viscosity stabilization
technique. Beside the usual conservation properties of mass, momentum and
total energy, the method also guarantees that the constructed solution
obeys pointwise stability constraints (positivity of density, internal
energy and a local minimum principle on the specific entropy). As such
\text{step-69} is strictly speaking more a collocation-type discretization
than a variational formulation, even though it is implemented with finite
elements.

The time-update at each node requires the evaluation of a right-hand side
that depends (nonlinearly) on information from the previous time-step that
spans more than one cell. Therefore, assembly loops operate directly on the
sparsity graph in order to retrieve information from the entire stencil
associated to each node. From a programming perspective, \texttt{step-69}
features a number of techniques that are of interest for a wider audience:
It discusses a hybrid thread and MPI parallelized scheme with efficient
(MPI node) local numbering of degrees of freedom. It showcases how to
perform asynchronous write-out of results using a background thread with
\texttt{std::async}, and discusses a simple but effective checkpointing and
restart technique.

\item \texttt{step-70}
  \todo[inline]{Luca, please write.}
\end{itemize}

\todo[inline]{Do we have new code gallery programs?}
\begin{itemize}
  \item Parallel in time (\url{https://github.com/dealii/code-gallery/pull/49})
  \item Linear elastoplasticity (\url{https://github.com/dealii/code-gallery/pull/62})
  \item Thermo-electro-elastic pump (\url{https://github.com/dealii/code-gallery/pull/63})
  \item Nonlinear poro-viscoelasticity (\url{https://github.com/dealii/code-gallery/pull/72})
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Python interfaces}
\label{subsec:python}

Initial support for Python has existed in \dealii{} since version
9.0. The present release significantly extends the Python
interface. Specifically, a large number of methods from classes such
as \texttt{Triangulation, CellAccessor, TriaAccessor, Mapping,
  Manifold, GridTools} can now be invoked from Python. We have focused
on methods and functions that are widely used when a mesh is created
and parameters related to the boundary, manifold, and material
identifiers are assigned. The following listing gives an idea of how
such code looks:
\begin{python}
import PyDealII.Release as dealii

triangulation = dealii.Triangulation('2D')
triangulation.generate_hyper_shell(center = dealii.Point([0, 0]),
                    inner_radius = 0.5, outer_radius = 1.,
                    n_cells = 0, colorize = True)

triangulation.refine_global(2)

for cell in triangulation.active_cells():
    cell.material_id = 1 if cell.center().x > 0 else 2

    for face in cell.faces():
        if face.at_boundary() and\
           face.boundary_id == 1:
            cell.refine_flag ='isotropic'

triangulation.execute_coarsening_and_refinement()
\end{python}
The mesh that results from this code is shown in Fig.~\ref{fig:pymesh}.

\begin{floatingfigure}[r]{0.3\textwidth}
  \centering
  \vspace*{-1.4cm}
  \includegraphics[width=0.3\textwidth]{python_mesh.png}
  \vspace*{-5mm}
  \caption{\it The mesh generated by the Python code shown in the main
    text. Cells are colored by material id.}
  \vspace*{0.2cm}
\label{fig:pymesh}
\end{floatingfigure}
All triangulations created from within Python are serial. However,
once the mesh is designed, the triangulation can be serialized along
with the auxiliary information about possible refinement, boundaries,
materials and manifolds. This object can be easily deserialized within
a C++ program for subsequent production runs. Furthermore, such a
serialized triangulation can also be used in the construction of
\texttt{parallel::shared}, \texttt{parallel::disrtibuted}, and
\texttt{parallel::fullydistributed} triangulations (see also Section~\ref{subsec:pft}).

To facilitate the illustration of the new Python bindings, tutorial programs \href{https://github.com/dealii/dealii/blob/dealii-9.2/examples/step-49/step-49.ipynb}{step-49} and \href{https://github.com/dealii/dealii/blob/dealii-9.2/examples/step-53/step-53.ipynb}{step-53} were replicated as Jupyter notebooks.

The introspective nature of the Python language makes it easy to infer
the list of supported methods from the Python objects, for example by
typing \texttt{dir(PyDealII.Release.Triangulation)}. 
The current Python interface does not yet provide access to \dealii{}'s finite element machinery, i.e., classes such as \texttt{DoFHandler, FE\_*, FEValues}, etc.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Incompatible changes}

The 9.2 release includes
\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_1_1_and_9_2_0.html}
     {around 60 incompatible changes}; see \cite{changes92}. The majority of these changes
should not be visible to typical user codes; some remove previously
deprecated classes and functions; and the majority change internal
interfaces that are not usually used in external
applications. That said, the following are worth mentioning since they
may have been more widely used:
\begin{itemize}
\item Two functions that provide information about all processes,
  namely
  \begin{itemize}
 \item \texttt{DoFHandler::locally\_owned\_dofs\_per\_processor()}
 \item \texttt{DoFHandler::locally\_owned\_mg\_dofs\_per\_processor()}
  \end{itemize}
have been deprecated. As discussed in
Subection~\ref{subsec:performance}, \dealii{} by default no longer
stores information for all processes on all processes, but only
local or locally-relevant information. On the other
hand, if necessary, global information can still be computed using,
for example, calling
\texttt{Utilities::MPI::Allgather(locally\_owned\_info(), comm)}.

\item A corresponding change has been made in parallel triangulation
  classes:
  \texttt{parallel::} \texttt{TriangulationBase::compute\_n\_locally\_owned\_active\_cells\_per\_processor()}
  can be used to obtain information about how many cells each
  process owns.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{How to cite \dealii}\label{sec:cite}

In order to justify the work the developers of \dealii{} put into this
software, we ask that papers using the library reference one of the
\dealii{} papers. This helps us justify the effort we put into it.

There are various ways to reference \dealii{}. To acknowledge the use of
the current version of the library, \textbf{please reference the present
document}. For up to date information and a bibtex entry
see
\begin{center}
 \url{https://www.dealii.org/publications.html}
\end{center}

The original \dealii{} paper containing an overview of its
architecture is \cite{BangerthHartmannKanschat2007}. If you rely on
specific features of the library, please consider citing any of the
following:
\begin{multicols}{2}
  \vspace*{-36pt}
  \begin{itemize}
 \item For geometric multigrid: \cite{Kanschat2004,JanssenKanschat2011,ClevengerHeisterKanschatKronbichler2019};
 \item For distributed parallel computing: \cite{BangerthBursteddeHeisterKronbichler11};
 \item For $hp$~adaptivity: \cite{BangerthKayserHerold2007};
  \item For partition-of-unity (PUM) and enrichment methods of the
    finite element space: \cite{Davydov2016};
 \item For matrix-free and fast assembly techniques:
   \cite{KronbichlerKormann2012,KronbichlerKormann2019};
 \item For computations on lower-dimensional manifolds:
   \cite{DeSimoneHeltaiManigrasso2009};
 \item For curved geometry representations and manifolds:
   \cite{HeltaiBangerthKronbichlerMola2019};
   \vfill\null  \columnbreak
 \item For integration with CAD files and tools:
   \cite{HeltaiMola2015};
 \item For boundary element computations:
   \cite{GiulianiMolaHeltai-2018-a};
 \item For \texttt{LinearOperator} and \texttt{PackagedOperation} facilities:
   \cite{MaierBardelloniHeltai-2016-a,MaierBardelloniHeltai-2016-b};
 \item For uses of the \texttt{WorkStream} interface:
   \cite{TKB16};
 \item For uses of the \texttt{ParameterAcceptor} concept, the
   \texttt{MeshWorker::ScratchData} base class, and the
   \texttt{ParsedConvergenceTable} class:
   \cite{SartoriGiulianiBardelloni-2018-a};
 \item For uses of the particle functionality in \dealii{}:
   \cite{GLHPB18}.
   \vfill\null  
\end{itemize}
\end{multicols}

\dealii{} can interface with many other libraries:
\todo[inline]{We picked up gingko. Anything else?}
\begin{multicols}{3}
\begin{itemize}
\item ADOL-C \cite{Griewank1996a,adol-c}
\item ARPACK \cite{arpack}
\item Assimp \cite{assimp}
\item BLAS and LAPACK \cite{lapack}
\item cuSOLVER \cite{cusolver}
\item cuSPARSE \cite{cusparse}
\item Gmsh \cite{geuzaine2009gmsh}
\item GSL \cite{gsl2016}
\item Ginkgo \cite{ginkgo-web-page}
\item HDF5 \cite{hdf5}
\item METIS \cite{karypis1998fast}
\item MUMPS \cite{ADE00,MUMPS:1,MUMPS:2,mumps-web-page}
\item muparser \cite{muparser-web-page}
\item nanoflann \cite{nanoflann}
  %
  % nanoflann has been deprecated and will be removed for 9.3. remove
  % this line in the 9.3 release paper
  %
\item NetCDF \cite{rew1990netcdf}
  %
  % netcdf has been deprecated and will be removed for 9.3. remove
  % this line in the 9.3 release paper
  %
\item OpenCASCADE \cite{opencascade-web-page}
\item p4est \cite{p4est}
\item PETSc \cite{petsc-user-ref,petsc-web-page}
\item ROL \cite{ridzal2014rapid}
\item ScaLAPACK \cite{slug}
\item SLEPc \cite{Hernandez:2005:SSF}
\item SUNDIALS \cite{sundials}
\item SymEngine \cite{symengine-web-page}
\item TBB \cite{Rei07}
\item Trilinos \cite{trilinos,trilinos-web-page}
\item UMFPACK \cite{umfpack}
\end{itemize}
\end{multicols}
Please consider citing the appropriate references if you use interfaces to these
libraries.

The two previous releases of \dealii{} can be cited as
\cite{dealII90,dealII91}.

\section{Acknowledgments}

\dealii{} is a world-wide project with dozens of contributors around the
globe. Other than the authors of this paper, the following people
contributed code to this release:\\
%
% Uwe Koecher doesn't usually show up in the changelog, but
% we should make sure he's listed.
%
% 9.2: updated 5/11/2020 MM
\todo[inline]{TODO: remove authors of paper}
Pasquale Africa,
Ashna Aggarwal,
Giovanni Alzetta,
Mathias Anselmann,
Daniel Arndt,
Wolfgang Bangerth,
Kirana Bergstrom,
Manaswinee Bezbaruah,
Bruno Blais,
Benjamin Brands,
Yong-Yong Cai,
Fabian Castelli,
Thomas C. Clevenger,
Katherine Cosburn,
Denis Davydov,
Elias Dejene,
Stefano Dominici,
Brett Dong,
Luel Emishaw,
Marc Fehling,
Niklas Fehn,
Rebecca Fildes,
Menno Fraters,
Andres Galindo,
Daniel Garcia-Sanchez,
Rene Gassmoeller,
Melanie Gerault,
Nicola Giuliani,
Brandon Gleeson,
Anne Glerum,
Krishnakumar Gopalakrishnan,
Graham Harper,
Mohammed Hassan,
Nicole Hayes,
Bang He,
Johannes Heinz,
Timo Heister,
Luca Heltai,
Jiuhua Hu,
Lise-Marie Imbert-Gerard,
Manu Jayadharan,
Daniel Jodlbauer,
Marie Kajan,
Guido Kanschat,
Alexander Knieps,
Uwe K{\"o}cher,
Martin Kronbichler,
Paras Kumar,
Konstantin Ladutenko,
Charu Lata,
Adam Lee,
Wenyu Lei,
Matthias Maier,
Katrin Mang,
Mae Markowski,
Franco Milicchio,
Adriana Morales Miranda,
Peter Munch,
Bob Myhill,
Emily Novak,
Omotayo Omosebi,
Alexey Ozeritskiy,
Jean-Paul Pelteret,
Rebecca Pereira,
Geneva Porter,
Laura Prieto Saavedra,
Reza Rastak,
Roland Richter,
Jonathan Robey,
Irabiel Romero,
Matthew Russell,
Tonatiuh Sanchez-Vizuet,
Natasha S. Sharma,
Doug Shi-Dong,
Konrad Simon,
Stephanie Sparks,
Sebastian Stark,
Simon Sticko,
Jan Philipp Thiele,
Jihuan Tian,
Ignacio Tomas,
Sara Tro,
Bruno Turcksin,
Ferdinand Vanmaele,
Zhuoran Wang,
David Wells,
Michal Wichrowski,
Julius Witte,
Winnifried Wollner,
Ming Yang,
Mario Zepeda Aguilar,
Wenjuan Zhang,
Victor Zheng.

Their contributions are much appreciated!

\bigskip

\dealii{} and its developers are financially supported through a
variety of funding sources:

D.~Arndt and B.~Turcksin: Research sponsored by the Laboratory Directed Research and
Development Program of Oak Ridge National Laboratory, managed by UT-Battelle,
LLC, for the U. S. Department of Energy.

W.~Bangerth, T.~C.~Clevenger, and T.~Heister were partially
supported by the National Science Foundation under award OAC-1835673
as part of the Cyberinfrastructure for Sustained Scientific Innovation (CSSI)
program  and by the Computational Infrastructure
in Geodynamics initiative (CIG), through the National Science
Foundation under Award No.~EAR-1550901 and The
University of California -- Davis.

W.~Bangerth and T.~Heister were also partially supported by award DMS-1821210.
\todo[inline]{Need to add FRES project}

%D.~Davydov was supported by the German Research Foundation (DFG), grant DA
%1664/2-1 and the Bayerisches Kompetenznetzwerk
%f\"ur Technisch-Wissenschaftliches Hoch- und H\"ochstleistungsrechnen
%(KONWIHR).

B.~Blais was partially supported by the National Science and Engineering Research Council of Canada(NSERC)  through the RGPIN-2020-04510 Discovery Grant

A.~V.~Grayver was partially supported by the European Space Agency
Swarm DISC program.

T.~Heister was also partially supported by NSF Award DMS-1522191, and
by Technical Data Analysis, Inc. through US Navy SBIR N16A-T003.

M.~Kronbichler was supported by the German
Research Foundation (DFG) under the project ``High-order discontinuous
Galerkin for the exa-scale'' (\mbox{ExaDG}) within the priority program ``Software
for Exascale Computing'' (SPPEXA) and the Bayerisches Kompetenznetzwerk
f\"ur Technisch-Wissen\-schaft\-li\-ches Hoch- und H\"ochstleistungsrechnen
(KONWIHR) in the context of the project
``Performance tuning of high-order discontinuous Galerkin solvers for
SuperMUC-NG''.

M.~Maier was partially supported by ARO MURI Award No. W911NF-14-0247 and
NSF Award DMS-1912847.

D.~Wells was supported by the National Science Foundation (NSF) through Grant
DMS-1344962.

Z.~Wang was partially
supported by the National Science Foundation under award OAC-1835673.

The Interdisciplinary Center for Scientific Computing (IWR) at Heidelberg
University has provided hosting services for the \dealii{} web page.

\bibliography{paper}{}
\bibliographystyle{abbrv}

\end{document}