### 2021-05-28 9.3/paper.tex
Go through the simplex section.

\documentclass{ansarticle-preprint}
%\usepackage{ucs}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
%\usepackage{cite}
\usepackage{anslistings}
\usepackage{multicol}
\usepackage{pdfsync}

\usepackage{pgfplots}
\usepackage{pgfplotstable}

\usepackage{fontenc}
\usepackage{graphicx}
\usepackage{xspace}

\usepackage{siunitx}

\usepackage{floatflt}

\usepackage{multirow}

%\renewcommand{\baselinestretch}{2.0}

\usepackage[normalem]{ulem}

\usepackage{todonotes}

\pgfplotsset{compat=1.9}
\definecolor{gnuplot@lightblue}{RGB}{87,181,232}
\definecolor{gnuplot@green}{RGB}{0,158,115}
\definecolor{gnuplot@purple}{RGB}{148,0,212}

\newcommand{\specialword}[1]{\texttt{#1}}
\newcommand{\dealii}{{\specialword{deal.II}}\xspace}
\newcommand{\pfrst}{{\specialword{p4est}}\xspace}
\newcommand{\trilinos}{{\specialword{Trilinos}}\xspace}
\newcommand{\aspect}{\specialword{Aspect}\xspace}
\newcommand{\petsc}{\specialword{PETSc}\xspace}
\newcommand{\cmake}{{\specialword{CMake}}\xspace}
\newcommand{\autoconf}{{\specialword{autoconf}}\xspace}

\usetikzlibrary{shapes.misc}
\tikzset{cross/.style={cross out, draw=black, minimum size=2*(#1-\pgflinewidth), inner sep=0pt, outer sep=0pt},
%default radius will be 1pt.
cross/.default={2pt}}

%
% Author list -- please add yourself in both places below (in
%                alphabetical order) if you think that your
%                contributions to the last release warrant this
%

\hypersetup{
  pdfauthor={
     Daniel Arndt,
     Wolfgang Bangerth,
    Marc Fehling,
    Timo Heister,
    Luca Heltai,
    Martin Kronbichler,
    Matthias Maier,
    Peter Munch,
    Jean-Paul Pelteret,
    Bruno Turcksin,
    David Wells
  },
  pdftitle={The deal.II Library, Version 9.3, 2021},
}

\title{The \dealii{} Library, Version 9.3}

 \author[1*]{Daniel Arndt}
 \affil[1]{Computational Engineering and Energy Sciences Group,
   Computational Sciences and Engineering Division,
   Oak Ridge National Laboratory, 1 Bethel Valley Rd.,
   TN 37831, USA.
   \texttt{arndtd/turcksinbr@ornl.gov}}

 \author[2,3]{Wolfgang~Bangerth}
 \affil[2]{Department of Mathematics, Colorado State University, Fort
   Collins, CO 80523-1874, USA.
   \texttt{bangerth/marc.fehling@colostate.edu}}
 \affil[3]{Department of Geosciences, Colorado State University, Fort
   Collins, CO 80523, USA.}

\author[2]{Marc~Fehling}

\author[4]{Timo~Heister}
 \affil[4]{School of Mathematical and Statistical Sciences,
   Clemson University,
   Clemson, SC, 29634, USA
   {\texttt{tcleven/heister@clemson.edu}}}

\author[5]{Luca~Heltai}
\affil[5]{SISSA,
   International School for Advanced Studies,
   Via Bonomea 265,
   34136, Trieste, Italy.
   {\texttt{luca.heltai@sissa.it}}}

 \author[6,7]{Martin~Kronbichler}
 \affil[6]{Institute for Computational Mechanics,
   Technical University of Munich,
   Boltzmannstr.~15, 85748 Garching, Germany.
   {\texttt{kronbichler/munch@lnm.mw.tum.de}}}
 \affil[7]{Department of Information Technology,
   Uppsala University,
   Box 337, 751\,05 Uppsala, Sweden.
   {\texttt{martin.kronbichler@it.uu.se}}}

\author[8]{Matthias~Maier}
\affil[8]{Department of Mathematics,
  Texas A\&M University,
  3368 TAMU,
  College Station, TX 77845, USA.
  {\texttt{maier@math.tamu.edu}}}

\author[6,9]{Peter Munch}
 \affil[10]{Institute of Materials Research, Materials Mechanics,
 Helmholtz-Zentrum Geesthacht,
 Max-Planck-Str. 1, 21502 Geesthacht, Germany.
   {\texttt{peter.muench@hzg.de}}}

\author[10]{Jean-Paul~Pelteret}
\affil[10]{Independent researcher.
{\texttt{jppelteret@gmail.com}}}

% \author[12]{Reza Rastak}
% \affil[12]{Department of Civil and Environmental Engineering,
%   Stanford University,
%   Stanford, CA 94305, USA.
%   {\texttt{rastak@stanford.edu}}}

% \author[**13]{Ignacio Tomas}
% \affil[13]{Sandia National Laboratories, Org 1442,
% P.O. Box 5800, MS 1320, Albuquerque, NM, 87185-1320,
%   \texttt{itomas@sandia.gov}}

\author[1*]{Bruno~Turcksin}
%

%\author[14]{Zhuoran Wang}
%\affil[14]{Department of Mathematics, Colorado State University, Fort Collins,
%          CO 80523-1874, USA.
%{\texttt{zhrwang@math.colostate.edu}}}

\author[15]{David Wells}
\affil[15]{Department of Mathematics, University of North Carolina,
  Chapel Hill, NC 27516, USA.
  {\texttt{drwells@email.unc.edu}}}

\renewcommand{\labelitemi}{--}

\begin{document}
\maketitle

\footnotetext{%
  $^\ast$ This manuscript has been authored by UT-Battelle, LLC under Contract No.
  DE-AC05-00OR22725 with the U.S. Department of Energy.
  %The United States
  %Government retains and the publisher, by accepting the article for
  %publication, acknowledges that the United States Government retains a
  %non-exclusive, paid-up, irrevocable, worldwide license to publish or reproduce
  %the published form of this manuscript, or allow others to do so, for United
  %States Government purposes. The Department of Energy will provide public
  %access to these results of federally sponsored research in accordance with the
  %DOE Public Access Plan (http://energy.gov/downloads/doe-public-access-plan).
}

\begin{abstract}
  This paper provides an overview of the new features of the finite element
  library \dealii, version 9.3.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Overview}

\dealii{} version 9.3.0 was released May ??, 2021.
\todo{Update date}
This paper provides an
overview of the new features of this release and serves as a citable
reference for the \dealii{} software library version 9.3. \dealii{} is an
object-oriented finite element library used around the world in the
development of finite element solvers. It is available for free under the
GNU Lesser General Public License (LGPL). Downloads are available at
\url{https://www.dealii.org/} and \url{https://github.com/dealii/dealii}.

The major changes of this release are:
%
\begin{itemize}
  \item Experimental support for simplex and mixed meshes (see Section~\ref{subsec:simplex});
  \item Improved flexibility of the particle infrastructure (see Section~\ref{subsec:particles});
  \item Support for global-coarsening multigrid algorithms (see Section~\ref{subsec:mg});
  \item Advances in the matrix-free infrastructure  (see Section~\ref{subsec:mf});
  \item Usage of MPI-3.0 shared-memory features to reduce memory footprint  (see Section~\ref{subsec:sm});
  \item Improved support for evaluation and integration at arbitrary points (see Section~\ref{subsec:fepointvalues});
  \item Nine new tutorial programs and a new code gallery program (see Section~\ref{subsec:steps}).
\end{itemize}
%
These major changes are discussed in detail in Section~\ref{sec:major}. There
are a number of other noteworthy changes in the current \dealii{} release
that we briefly outline in the remainder of this section:
%
\begin{itemize}
  \item Each non-artificial cell now has a globally unique index that can be queried
  for active cells via \texttt{CellAccessor::global\_active\_cell\_index()} and for level cells 
  via \texttt{::global\_\allowbreak level\_\allowbreak cell\_\allowbreak index()}. The information
  can be used to efficiently index into global vectors storing data
  for each cell, rather than for each degree of freedom corresponding to a finite
  element field.

  \item Previously, functions and classes were marked using the 
  macro \texttt{DEAL\_II\_DEPRECATED} and then typically removed in
  the release after the one in which these deprecation notices were
  available to users. We have now extended this policy by introducing
  the \texttt{DEAL\_II\_\allowbreak DEPRECATED\_EARLY} macro, which indicates that a feature will be
  deprecated in the next release. In contrast to the first macro, it will only give
  warnings if \dealii{} has been configured with \texttt{-D DEAL\_II\_EARLY\_DEPRECATIONS=ON}.
  \item After each update of the master branch of \dealii{}, we build a new Docker image
  with all features enabled. It is
  accessible on \texttt{Docker Hub} via \texttt{dealii/dealii:master-focal}. This image is
  particular useful when used in the contiguous-integration processes of user codes.
\end{itemize}
%
The changelog lists more than 200 other features and bugfixes.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Major changes to the library}
\label{sec:major}

This release of \dealii{} contains a number of large and significant changes
that will be discussed in this section.
It of course also contains a
vast number of smaller changes and added functionality; the details of these
can be found
\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_2_0_and_9_3_0.html}
{in the file that lists all changes for this release}; see \cite{changes93}.

%\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Experimental simplex and mixed mesh support}
\label{subsec:simplex}

The current release of \dealii adds experimental support for simplex
meshes (consisting of triangles in 2D; tetrahedra in 3D) and mixed meshes (consisting of triangles and/or
quadrilaterals in 2D; tetrahedra, pyramids, wedges, and/or hexahedra in 3D).
Many freely available mesh-generation tools produce such kind of
meshes and they are widely used in industry and applications, but were
previously unsupported by \dealii. As a consequence, users of \dealii
had to pre-process such meshes and convert them to pure quadrilateral
or hexahedral meshes.

Support for simplex and mixed meshes is not universal in \dealii{} at
this point. While \dealii{} can read such meshes, write output for
them, and solve partial differential equations with certain finite
elements, there are also many areas that have not been fully converted
to the new functionality. In particular, \dealii{} currently only
offers low-order finite elements on such meshes, and many tool
functions might throw exceptions when used with such meshes.

A user-focused summary of information around simplex and mixed mesh
support is also available on the new module page at
  \url{https://www.dealii.org/current/doxygen/deal.II/group__simplex.html}
In particular, it shows how to solve a simple Poisson problem like in
the \texttt{step-3} tutorial program
on simplex and mixed meshes, with a focus on the necessary changes to
the workflow.
At the time of this release, there are also 92 tests (in the folder \texttt{tests/simplex})
targeting the new simplex and mixed mesh support. In particular, the folder also
contains ported variants of a number of existing tutorials: 1, 2, 3, 4, 6, 7, 8, 12, 17, 18, 20, 23, 31, 38,
40, 55, 67, 68, and 74.

\begin{figure}

  \centering

  \begin{tikzpicture}[scale=1.7]

    \coordinate (0) at (0,0);
    \coordinate (1) at (1,0);
    \coordinate (2) at (0,1);

    \coordinate (3) at (0,0.5);
    \coordinate (4) at (0.5,0.5);
    \coordinate (5) at (0.5,0);

    \foreach \i in {0,1, 2}
    \draw[blue,fill=blue] (\i) circle (1.1pt) node [below] {};

    \foreach \i in {3,4,5}
    \draw (\i) node[cross, draw=red, rotate=-20] {};

    \draw (0) --(1) -- (2) -- (0);

  \end{tikzpicture}
  \begin{tikzpicture}[scale=1.7]

    \coordinate (0) at (0,0);
    \coordinate (1) at (1,0);
    \coordinate (2) at (1,1);
    \coordinate (3) at (0,1);

    \coordinate (4) at (0.5,0,0);
    \coordinate (5) at (1.0,0.5);
    \coordinate (6) at (0.0,0.5);
    \coordinate (7) at (0.5,1.0);
    \coordinate (8) at (0.5,0.5);

    \foreach \i in {0,1, 2, 3}
    \draw[blue,fill=blue] (\i) circle (1.1pt) node [below] {};

    \foreach \i in {4,5,6,7,8}
    \draw (\i) node[cross, draw=red, rotate=-20] {};

    \draw (0) --(1) -- (2) -- (3) -- (0);

  \end{tikzpicture}
  \qquad\qquad
  \begin{tikzpicture}[scale=1.3]

    \coordinate (0) at (0,0,0);
    \coordinate (1) at (1,0,0);
    \coordinate (2) at (0,1,0);
    \coordinate (3) at (0,0,1);

    \coordinate (4) at (0.5,0,0);
    \coordinate (5) at (0,0.5,0);
    \coordinate (6) at (0,0,0.5);
    \coordinate (7) at (0.5,0.5,0.0);
    \coordinate (8) at (0.5,0.0,0.5);
    \coordinate (9) at (0.0,0.5,0.5);

    \foreach \i in {0,1,2,3}
    \draw[blue,fill=blue] (\i) circle (1.5pt) node [below] {};

    \foreach \i in {4,5,6,7,8,9}
    \draw (\i) node[cross, draw=red, rotate=-20] {};

    \draw (0) -- (1) -- (2) -- (0);
    \draw (0) -- (1) -- (3) -- (0);
    \draw (1) -- (2) -- (3) -- (1);

  \end{tikzpicture}
  \begin{tikzpicture}[scale=1.5]

    \coordinate (0) at (0.0, 0.0, 0.0);
    \coordinate (1) at (0.0, 0.0, 1.0);
    \coordinate (2) at (1.0, 0.0, 0.0);
    \coordinate (3) at (1.0, 0.0, 1.0);
    \coordinate (4) at (0.5, 1.0, 0.5);

    \coordinate  (5) at (0.5, 0.0, 0.0);
    \coordinate  (6) at (0.0, 0.0, 0.5);
    \coordinate  (7) at (0.5, 0.0, 1.0);
    \coordinate  (8) at (1.0, 0.0, 0.5);

    \coordinate  (9) at (0.25,0.5,0.25);
    \coordinate (10) at (0.25,0.5,0.75);
    \coordinate (11) at (0.75,0.5,0.25);
    \coordinate (12) at (0.75,0.5,0.75);

    \coordinate (13) at (0.5, 0.0, 0.5);

    \foreach \i in {0,1,2,3,4}
    \draw[blue,fill=blue] (\i) circle (1.3pt) node [below] {};

    %\foreach \i in {5, 6, 7, 8, 9, 10, 11, 12, 13}
    %    \draw (\i) node[cross] {};

    \draw (0) -- (1) -- (3) -- (2) -- (0);
    \draw (1) -- (4) -- (3);
    \draw (0) -- (4) -- (2);

  \end{tikzpicture}
  \begin{tikzpicture}[scale=1.3]

    \coordinate (0) at (0,0,1.5);
    \coordinate (1) at (1,0,1.5);
    \coordinate (2) at (0.0,1.0,1.5);

    \coordinate (3) at (0,0,0);
    \coordinate (4) at (1,0,0);
    \coordinate (5) at (0.0,1.0,0);

    \coordinate  (6) at (0.5, 0.0, 0.0);
    \coordinate  (7) at (0.0, 0.5, 0.0);
    \coordinate  (8) at (0.5, 0.5, 0.0);

    \coordinate   (9) at (0.5, 0.0, 0.75);
    \coordinate  (10) at (0.0, 0.5, 0.75);
    \coordinate  (11) at (0.5, 0.5, 0.75);

    \coordinate  (12) at (0.5, 0.0, 1.5);
    \coordinate  (13) at (0.0, 0.5, 1.5);
    \coordinate  (14) at (0.5, 0.5, 1.5);

    \foreach \i in {0,1,2,3,4,5}
    \draw[blue,fill=blue] (\i) circle (1.5pt) node [below] {};

    \foreach \i in {6,7,8,9,10,11,12,13,14}
    \draw (\i) node[cross, draw=red, rotate=-20] {};

    \draw (0) -- (1) -- (2) -- (0);
    \draw (3) -- (4) -- (5) -- (3);
    \draw (0) -- (3);
    \draw (1) -- (4);
    \draw (2) -- (5);

  \end{tikzpicture}
  \begin{tikzpicture}[scale=1.5]

    \coordinate (0) at (0,0,1);
    \coordinate (1) at (1,0,1);
    \coordinate (2) at (0,0,0);
    \coordinate (3) at (1,0,0);

    \coordinate (4) at (0,1,1);
    \coordinate (5) at (1,1,1);
    \coordinate (6) at (0,1,0);
    \coordinate (7) at (1,1,0);

    \coordinate  (8) at (0,0,0.5);
    \coordinate  (9) at (1,0,0.5);
    \coordinate (10) at (0.5,0,1);
    \coordinate (11) at (0.5,0,0);
    \coordinate (12) at (0.5,0,0.5);

    \coordinate (13) at (0,1,0.5);
    \coordinate (14) at (1,1,0.5);
    \coordinate (15) at (0.5,1,1);
    \coordinate (16) at (0.5,1,0);
    \coordinate (17) at (0.5,1,0.5);

    \coordinate (18) at (0,0.5,1);
    \coordinate (19) at (1,0.5,1);
    \coordinate (20) at (0,0.5,0);
    \coordinate (21) at (1,0.5,0);
    \coordinate (22) at (0,0.5,0.5);
    \coordinate (23) at (1,0.5,0.5);
    \coordinate (24) at (0.5,0.5,1);
    \coordinate (25) at (0.5,0.5,0);
    \coordinate (26) at (0.5,0.5,0.5);

    \foreach \i in {0,1,2,3,4,5,6,7}
    \draw[blue,fill=blue] (\i) circle (1.3pt) node [below] {};

    \foreach \i in {8,9,10,11,12,13,14,15,16,17,18,19,20,21,22,23,24,25,26}
    \draw (\i) node[cross, draw=red, rotate=-20] {};

    \draw (0) -- (1) -- (3) -- (2) -- (0);
    \draw (4) -- (5) -- (7) -- (6) -- (4);
    \draw (0) -- (4);
    \draw (1) -- (5);
    \draw (2) -- (6);
    \draw (3) -- (7);

  \end{tikzpicture}

  \caption{\it Reference cells in 2D (triangle, quadrilateral) and 3D
    (tetrahedron, pyramid, wedge, hexahedron) with the support points
    indicated both for linear ({\color{blue}$\bullet$}) and quadratic
    ({\color{red}$\times$}) shape functions.
  \todo[inline]{The pyramid is missing red crosses. The wedge is
    missing them on three edges.}}
  \label{fig:simplex}
\end{figure}

\begin{table}
  \caption{\it List of new scalar \texttt{FiniteElement} classes for the new
  reference-cell types. Vectorial elements can be constructed based on these classes
  via \texttt{FE\_Systems}.}\label{tab:simplex:fe}
  \centering
  \begin{tabular}{cccc}
    \toprule
    \textbf{reference cell} & \textbf{finite element} & \textbf{dimension} & \textbf{degree}\\
    \midrule
    \multirow{2}{*}{simplex (line, triangle, tetrahedron)} & FE\_SimplexP, FE\_SimplexDGP  & 1--3 & 1--2 \\
                                          & FE\_SimplexP\_Bubbles  & 1--3 & 1--2 \\
    pyramid & FE\_PyramidP, FE\_PyramidDGP  & 3 & 1 \\
    wedge & FE\_WedgeP, FE\_WedgeDGP  & 3 & 1--2 \\
    \bottomrule
  \end{tabular}
\end{table}

\subsubsection{Refactoring of internal data structures}

To enable simplex and mixed mesh support, we performed a major refactoring of the internal data structures of \dealii. In particular, the \texttt{Triangulation} class and
the \texttt{DoFHandler} class have undergone large changes and now
support meshes composed of all of the cells shown in Fig.~\ref{fig:simplex}.

The template parameters of the internal data structures of \texttt{Triangulation}
have been removed and the type of each cell and of each face (only in 3D) is stored. The function
\texttt{Triangulation::create\_\allowbreak triangulation()}, which converts a given list of
cells and vertices to the internal data structures, has been rewritten
inspired by \citep{logg2012} and this has had the side effect of a speed-up of up to 5. Minor adjustments
have also been made to the \texttt{parallel::shared::Triangulation}
and \texttt{parallel::\allowbreak fullydistributed::\allowbreak
  Triangulation} classes such that
the new mesh types can also be used in parallel.

The internal data structures of \texttt{DoFHandler} used to be hard-coded for pure
hypercube meshes, while the \texttt{hp::DoFHandler} used to be built around
CRS-like data structures. Due to the need of CRS data structures in the
\texttt{DoFHandler} in the context of more general meshes, we have merged
\texttt{hp::DoFHandler} into the \texttt{DoFHandler}. The class \texttt{hp::DoFHandler}, which is now a dummy derivation of \texttt{DoFHandler}, currently only exists for compatibility reasons and has
been deprecated, see Section~\ref{subsec:deprecated}. It will be
removed in the next release.

\subsubsection{Generating meshes}

The most obvious way to generate a simplex or a mixed mesh is to read the mesh from a file
generated by an external mesh generator. Currently, we support the
\texttt{VTK}, \texttt{MSH} (generated by the \textsc{gmsh} program \cite{geuzaine2009gmsh}), and \texttt{EXODUS II} file formats.

Alternatively, one can create a pure hypercube mesh with the existing functions
in the \texttt{GridGenerator} namespace and convert it to a
pure simplex mesh with the function
\texttt{GridGenerator::convert\_\allowbreak hypercube\_\allowbreak to\_\allowbreak simplex\_\allowbreak mesh()}.

\subsubsection{Using simplex meshes}

After having created a triangulation, one can proceed in the same way
as for hypercube meshes. In particular, one selects an appropriate finite element, mapping, and
quadrature class as follows:

\begin{c++}
FE_SimplexP<dim, spacedim> fe(degree);
MappingFE<dim, spacedim> mapping(FE_SimplexP<dim, spacedim>(1));
QGaussSimplex<dim> quad(degree + 1);

DoFHandler<dim> dof_handler(tria);
dof_handler.distribute_dofs(fe);

FEValues<dim, spacedim> fe_values(mapping, fe, quad, flags);
\end{c++}
The  list of currently supported finite-element classes is provided in Table~\ref{tab:simplex:fe}. Currently,
only linear iso-parametric mappings (via the \texttt{MappingFE} and
\texttt{MappingFEField} classes) are available.
\todo[inline]{I don't understand the previous sentence. "linear" means
  p=1, but isoparametric means that it has the same p as the finite
  element. It can't be both.}
For quadrature, the classes \texttt{QDuffy}, \texttt{QGaussSimplex}, \texttt{QWitherdenVincentSimplex},
\texttt{QGaussPyamid}, and \texttt{QGaussWedge} are
available.

\subsubsection{Using mixed meshes}
For mixed meshes, concepts known from the $hp$-context have been applied:
 different finite-element classes are assigned to different cells
 based on their respective kind of reference cell. In the
case of a 2D mixed mesh, which can only consist of triangles and
quadrilaterals, the finite element defined on a triangle (e.g., \texttt{FE\_SimplexP})
and on a quadrilateral (e.g., \texttt{FE\_Q}) can be collected in a \texttt{hp::FECollection}:

\begin{c++}
hp::FECollection<dim, spacedim> fe
  {FE_SimplexP<dim, spacedim>(degree), FE_Q<dim, spacedim>(degree)};
\end{c++}

Similarly, \texttt{hp::QCollection} and \texttt{hp::MappingCollection} can be used
to construct appropriate collections. Furthermore, the correct active
finite element index, which points to the correct
finite element of that cell, has to be assigned to each cell. The
following piece of code will then correctly enumerate all degrees of
freedom on the mesh:

\begin{c++}
DoFHandler<dim> dof_handler(tria);

for (const auto &cell : dof_handler.active_cell_iterators())
  switch(cell->reference_cell_type())
    {
      case ReferenceCell::Type::Tri:  cell->set_active_fe_index(0); break;
      case ReferenceCell::Type::Quad: cell->set_active_fe_index(1); break;
      // 3D (Tet, Pyramid, Wedge, Hex) not shown
      default: Assert(false, ExcNotImplemented());
    }

dof_handler.distribute_dofs(fe);
\end{c++}

\subsubsection{Practical implications}

The introduction of simplex and mixed meshes leads to some implications
for the user if these features are to be used. For instance, each cell might have a different type with
different number of vertices, lines, and faces so that these quantities can not be
compile-time constants anymore. This information used to be queried from
the \texttt{GeometryInfo} class. To avoid using this class, we have extended
relevant classes, e.g., \texttt{TriaAccessor} or \texttt{TriaCellAccessor},
with useful new functions like \texttt{n\_vertices()}, \texttt{n\_lines()}, or
\texttt{n\_faces()} so that users can simply write:
\begin{c++}
for(const auto & cell : tria.active_cell_iterators())
  for(unsigned int f = 0; f < cell->n_faces(); ++f)
    // do something with cell->face(f);
\end{c++}
Alternatively, one can use an iterator-based approach to loop over all faces
of a cell, introduced in the previous release. The relevant functions have been adjusted to be able to deal with
the now variable number of faces per cell:
\begin{c++}
for(const auto & cell : tria.active_cell_iterators())
  for(const auto & face : cell->face_iterators())
    // do something with face
\end{c++}
Furthermore, for mixed meshes, the number of degrees of freedom will differ between cells so that
cell-local arrays need to be resized for each
cell (as has previously already been the case in the $hp$-context):
\begin{c++}
std::vector<double> local_rhs;
for(const auto & cell : tria.active_cell_iterators())
 {
   hp_fe_values.reinit(cell);
   local_rhs.resize(cell->get_fe().dofs_per_cell);
 }
\end{c++}

What is true for cells is also true for faces in 3D: faces can be either triangles
or quadrilaterals. This is the case even if no mixed mesh is used if
the mesh consists exclusively of pyramids or wedges. As a consequence, some functions, e.g. \texttt{FiniteElementData::n\_dofs\_per\_face()}, have been extended with a new optional argument for the face number.

Geometric information about cells and faces -- previously provided by
the \texttt{GeometryInfo} class for hypercube-type cells -- is now
available via the  \texttt{ReferenceCell} class that can be
instantiated for each of the seven possible reference cell. The
correct reference cell object for a cell or face can be obtained, respectively, using
\texttt{cell->reference\_cell()}, and either the call
\texttt{cell->reference\_cell().face\_reference\_cell(f)}
or
\texttt{cell->face(f)->reference\_cell()}.

Furthermore, many functions in \dealii{} used mappings that, when not
given explicitly, defaulted to (bi-/tri-)linear ones. These no longer
work for simplex or mixed meshes, and in these cases users will need
to explicitly provide the correct mapping for the mesh to be used.

\subsubsection{Matrix-free support}

\dealii{}'s matrix-free support has also been extended to simplex and
mixed meshes, for both
continuous and discontinuous elements. From a user perspective, the main
changes are to pass a $d$-dimensional quadrature object (rather than
one for 1D),
and the fact that \texttt{FEEvaluation} and \texttt{FEFaceEvaluation} must not
specify the polynomial degree via template arguments, determining all
information at runtime. More details can be found in
Section~\ref{subsec:mf}.

For the current release, no advanced algorithms for evaluating values and
gradients at quadrature points, such as sum factorization, are used.
The use of full interpolation matrices is for now acceptable since only low-order elements are
supported.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Advances in the particle infrastructure}
\label{subsec:particles}

TODO

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Advances in the multigrid infrastructure}
\label{subsec:mg}

Until now, \dealii has only supported local smoothing multigrid
algorithms~\citep{ClevengerHeisterKanschatKronbichler2019}; in local
smoothing algorithms, smoothers only act on the cells of a given refinement level,
skipping those parts of the mesh that are not adaptively refined to that level.
The current release now also has support for global coarsening~\citep{becker2000multigrid, sundar2012parallel} when
using continuous (\texttt{FE\_Q}, \texttt{FE\_SimplexP}) and
discontinuous (\texttt{FE\_DGQ}, \texttt{FE\_SimplexDGP})
elements. Global coarsening builds multigrid levels for the entire
domain, coarsening \textit{all} cells of a triangulation regardless of
how many times they have been refined. In addition, the framework now available in
\dealii{} in not only applicable to geometric coarsening, but can also
perform coarsening by reducing the polynomial degree ($p$-multigrid,
see~\cite{Fehn2020}), for example
to support $hp$-adaptive meshes.
%where it has been shown that one can
%efficiently precondition high-order spaces with lower-order finite
%element operators.
% MK: I do not think "precondition" is a precise enough description of p-multigrid
% because one still needs smoothing on the fine levels in addition to the
% "solve" at the coarser degree, a fact not transported by the mere
% 'precondition' term; rather than trying to discuss this in a more precise way,
% let us simply skip it as that is not central to the presentation here.
Finally, the implementation also supports transfer between continuous
and discontinuous elements as a further way to create multigrid levels.

These new multigrid variants promise fewer solver
iterations and better parallel scalability than the existing local
smoothing algorithms, but have to deal with
hanging nodes within each level and general require more computational
work per iteration overall.

The transfer operators between two levels have been implemented in the new class \texttt{MGTwoLevel\allowbreak Transfer}, which can be set up via the functions \texttt{MGTwoLevel\allowbreak Transfer::\allowbreak reinit\_\allowbreak geometric\_\allowbreak transfer()} or \texttt{MGTwo\allowbreak LevelTransfer::\allowbreak reinit\_\allowbreak polynomial\_\allowbreak transfer()} for given
\texttt{DoFHandler} and \texttt{AffineConstraint} classes of two levels. The resulting transfer operators
can then be collected in a single
\texttt{MGTransfer\allowbreak GlobalCoarsening} object that can be used just as the previous workhorse \texttt{MGTransferMatrixFree} within the \texttt{Multigrid}
algorithm.
%Several common operations are encoded in utility
%functions in the \texttt{MGTransferGlobalCoarseningTools}
%namespace.
To facilitate the construction of matrix diagonals with matrix-free methods as well as a matrix representation of the coarse level matrix, new utility functions \texttt{create\_diagonal()} and \texttt{create\_matrix()} from
the \texttt{MatrixFreeTools} namespace have been added (see also Subsection~\ref{subsec:mf}).

The usage of the new transfer operators (and of some of the utility
functions) in the context of a hybrid multigrid algorithm
($hp$-multigrid with algebraic multigrid as coarse-grid solver) for $hp$-problems is demonstrated in the new tutorial \texttt{step-75}, see also Section~\ref{subsec:steps}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Advances in the matrix-free infrastructure}
\label{subsec:mf}

\subsubsection{Precompilation of evaluation kernels}

The classes \texttt{FEEvaluation} and \texttt{FEFaceEvaluation} use template parameters for
the polynomial degree of the finite element $k$ and the number of the 1D quadrature points $q$
to generate near-optimal code for these operations.
For application codes that rely on operators of many different degrees (e.g., because
they use $p$-multigrid or $hp$-algorithms), creating all instantiations
can be overly complex and incur long compile times.

In the current release, specializations of these classes that do
not rely on the template parameters $k$ and $q$ (expressed in the code
with ``-1'' and ``0'') have been added. For example:
\begin{c++}
FEEvaluation<dim, -1, 0, n_components, Number, VectorizedArrayType>
  phi(range, dofhandler_index, quadrature_index, first_selected_component);
\end{c++}
These classes select at runtime -- for common low and medium
polynomial degree/quadrature combinations ($k\le 6$ and $q\in\{ k+1,
k+2, \left\lfloor (3k)/2 \right\rfloor \}$) -- efficient precompiled implementations and default to non-templated
evaluation kernels else.

In the case that even higher polynomial degrees are needed, one can precompile the
relevant internal classes
(\texttt{FEEvaluationFactory}, \texttt{FEFaceEvaluationFactory}, \texttt{CellwiseInverseMassFactory}) in the user code for needed \texttt{degree}s
and \texttt{VectorizedArrayType}s in the following way:
\begin{c++}
#define FE_EVAL_FACTORY_DEGREE_MAX 12

#include <deal.II/matrix_free/evaluation_template_factory.templates.h>

DEAL_II_NAMESPACE_OPEN
template struct dealii::internal::FEEvaluationFactory
  <dim, VectorizedArrayType::value_type, VectorizedArrayType>;
DEAL_II_NAMESPACE_CLOSE
\end{c++}

\subsubsection{Parallel matrix-free $hp$-implementation}\label{subsubsection:mf:hp}

With release 9.1 \cite{dealII91}, large parts of the $hp$-algorithms in \dealii were parallelized, enabling
parallel matrix-based simulations with the $hp$-infrastructure. In the present
release, the parallel $hp$ support was extended to \texttt{MatrixFree}.

Until now, the \texttt{FEEvaluation} classes used the template parameters $k$ and $q$ to select the correct active FE and quadrature index and cumbersome detection of subranges of the same degree.
The creation of subranges is now performed internally, and the non-templated versions
of the \texttt{FEEvaluation} classes have been extended for the $hp$-case. To determine
the desired FE and quadrature index of a subrange, the current cell/face range has to be provided
to the constructors of the \texttt{FEEvaluation} classes. These changes enable
users to write matrix-free code independently of whether $hp$-capabilities are used or not.

The new tutorial \texttt{step-75} presents how to use the new $hp$-related features in \texttt{MatrixFree}
in the context of a hybrid-multigrid solver.

\subsection{MPI-3.0 shared-memory support}\label{subsec:sm}
In many large computations, certain pieces of data are computed once
and then treated as read-only. If this information is needed by more
than one MPI process, it is more efficient to store this information only
once in shared memory among all processes located on a multicore
node. MPI supports the creation of such shared memory windows since
version 3.0, and deal.II can now use this in the
\texttt{AlignedVector} and \texttt{Table} classes that are often used
for large lookup tables.

The feature is also used in the \texttt{MatrixFree} and
\texttt{LinearAlgebra::\allowbreak distributed::\allowbreak Vector} classes. If \texttt{MatrixFree} has been configured by setting
\texttt{MatrixFree::AdditionalData::communicator\_sm} appropriately, then
\texttt{MatrixFree::create\_dof\_vector()} creates vectors that share
information among all processes on one node. As a consequence, the
\texttt{FEEvaluation} classes can access vector elements owned by
other processes and in certain cases node-local
communication can be skipped. To prevent race conditions, \texttt{MatrixFree} uses local
barriers at the beginning and the end of loops (\texttt{loop()}, \texttt{cell\_loop()}, \texttt{loop\_cell\_centric()}).

The new \texttt{step-76} tutorial program illustrates this latter case
in the context of the solution of the Euler equations. \texttt{step-76} reaches a
speed-up of 27\% compared to the
original version, \texttt{step-67}, by using the new feature.
For more details and the usage of the feature in the library \texttt{hyper.deal}, see \citep{munch2020hyperdeal}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Evaluation and integration at arbitrary points}
\label{subsec:fepointvalues}

In a number of circumstances, finite element solutions need to be evaluated on
arbitrary reference points that change from one element to next. Two important
examples are particle simulations coupled to a finite element solution, or
algorithms on non-matching grids. The existing \texttt{FEValues} class is a poor
fit for this task, as it relies on tabulating all information before its usage,
necessitating a separate setup for every cell. The new class
\texttt{FEPointEvaluation} provides a more convenient interface for this
task. For tensor product finite elements (\texttt{FE\_Q}, \texttt{FE\_DGQ})
and tensor product mappings (\texttt{MappingQGeneric} and derived classes),
the new approach is also very fast, as it can use some of the matrix-free
infrastructure and vectorization facilities.

To give an example of the new features, let us consider the evaluation of a
surface tension force in the context of sharp-interface methods, whose
contribution is added to a fluid solver by multiplication with test function
and addition over quadrature points:
\begin{align*}
\left(\vec{v}, \kappa \, \vec{n}\right)_\Gamma
\approx
\sum_q\left(\vec{v}, \kappa(\vec{p}_q) \, \vec{n}(\vec{p}_q) (JxW)_q\right).
\end{align*}
In \dealii{}, this can now be conveniently written as
\begin{c++}
phi_normal.reinit(cell, reference_points);
phi_force.reinit(cell, reference_points);
phi_normal.evaluate(normal_values, EvaluationFlags::values);
for (unsigned int q = 0; q < n_points; ++q)
   phi_force.submit_value(phi_curvature.get_value(q) *
                          phi_normal.get_value(q) * JxW[q], q);
phi_force.integrate(force_values, EvaluationFlags::values);
\end{c++}
The quadrature points (at reference positions - \texttt{reference\_points}) and the related \texttt{JxW} value can, e.g., come from
a mesh of codimension one. Determining to which \texttt{cell} a quadrature
point belongs to on the background mesh, including the reference-cell
coordinates \texttt{reference\_points}, can be determined with functions like
\texttt{GridTools::find\_all\_active\_cells\_around\_point()}. While these
functions have been available in \dealii{} previously, their performance has
been considerably enhanced with the aforementioned more optimized code paths
for selected mappings.

While \texttt{FEPointEvaluation} assumes that evaluation points are already sorted according to 
the owning cells and such can concentrate on cell-local operations, the new class
\texttt{RemovePointEvaluation} is responsible to determine the owning cells in a distributed
context and for providing efficient communication patterns for the data exchange. In \texttt{deal.II},
the class has been successfully applied together with \texttt{FEPointEvaluation} to evaluate a distributed
solution vector at arbitrary points (see \texttt{VectorTools::evaluate\_at\_points()}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{New and improved tutorials and code gallery programs}
\label{subsec:steps}

Many of the \dealii{} tutorial programs were revised in a variety of
ways as part of this release. TODO

In addition, there are a number of new tutorial programs:
\begin{itemize}
\item \texttt{step-19} is an introductory demonstration of \dealii{}'s
  particle functionality. It solves the coupled problem of
  charged particles and an electric field, using a cathode tube as an
  example.

\item \texttt{step-68} TODO

\item \texttt{step-71} focuses on automatic and symbolic
  differentiation (AD and SD, in short) as a tool to make solvers for complex,
  nonlinear problems possible. To this end, \dealii{} can interface to a number
  of AD and SD libraries, specifically Trilinos' Sacado package
  \cite{Bartlett2006a}, ADOL-C \cite{Griewank1996a}, and
  SymEngine \cite{symengine-web-page}. The tutorial
  illustrates how these techniques can be used to compute derivatives
  of first a rather simple function, and then of the much more complex
  energy functions of two magnetoelastic and magneto-viscoelastic
  material formulations in which just the scalar energy functional
  takes up the better part of a page, and even first derivatives can
  only be computed with heroic effort.

\item \texttt{step-72} illustrates the use of automatic
  differentiation to simplify the computation of derivatives in the
  context of nonlinear partial differential equations where one needs
  to compute the Jacobian from the residual operator for efficient
  Newton iterations. \texttt{step-72} builds on the minimal surface
  solver \texttt{step-15} and replaces the hand-construction of the
  Jacobian by either computing it as the derivative of the residual,
  or alternatively as the second derivative of an energy functional
  that then also yields the residual itself.

\item \texttt{step-75} demonstrates a state-of-the-art way of solving a simple
      Laplace problem using $hp$-adaptation and hybrid multigrid methods on machines
      with distributed memory. This tutorial points out particularities in porting
      serial $hp$-adaptive code for parallelization. Furthermore, it guides through
      the process of writing an efficient preconditioner for $hp$-adaptive applications.

\item \texttt{step-76} is an explicit time integrator for the
      compressible Euler equations discretized with a high-order discontinuous
      Galerkin (DG) scheme, using the matrix-free infrastructure just as \texttt{step-67} does.
      The tutorial presents advanced topics, like the usage of cell-centric loops and
      the new MPI-3.0 shared-memory capabilities of \texttt{MatrixFree} tor each high
      throughput. Furthermore, the utilization of the template parameter
      \texttt{VectorizedArrayType} and the application of lambdas to capture cell and face
      integrals are discussed.

\item \texttt{step-77} is a program that illustrate \texttt{dealii}'s
  interfaces to the SUNDIALS library \cite{sundials}, and specifically
  the KINSOL nonlinear solver. Like the \texttt{step-72} program
  mentioned above, it is a variation of the minimal surface solver
  \texttt{step-15}: Instead of implementing the nonlinear Newton and
  line search loop ourselves, \texttt{step-77} relies on KINSOL for
  decisions such as when to rebuild the Jacobian matrix, when to
  actually solve linear systems with it, and how to form updates that
  drive the residual to convergence. The program illustrates the
  substantial savings that can be obtained by not re-inventing the
  wheel but instead building on an existing and well-tuned software
  such as KINSOL.

  \dealii{}'s interfaces were also updated to the latest SUNDIALS release, 5.7.

\item \texttt{step-78} TODO

\item \texttt{step-79} TODO
\end{itemize}

There is also a new program in the code gallery (a collection of
user-contributed programs that often solve more complicated problems
than tutorial programs, and intended as starting points for further
research rather than as teaching tools):
\begin{itemize}
  \item preCICE: TODO
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Incompatible changes}\label{subsec:deprecated}

The 9.3 release includes
\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_2_0_and_9_3_0.html}
{around 45 incompatible changes}; see \cite{changes93}. The majority of these changes
should not be visible to typical user codes; some remove previously
deprecated classes and functions; and the majority change internal
interfaces that are not usually used in external
applications. That said, the following are worth mentioning since they
may have been more widely used:
\begin{itemize}
  \item The \texttt{hp::DoFHandler} class has been deprecated. The standard
        \texttt{DoFHandler} class is now capable of all $hp$-functionalities.
  \item Consequently, all template arguments \texttt{DoFHandlerType} are obsolete,
        and classes like \texttt{DataOut} or \texttt{SolutionTransfer} for example
        no longer require them. If you rely on these template arguments, an interim
        namespace \texttt{Legacy} has been introduced that provides all affected
        classes with the old interface for the transition phase.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{How to cite \dealii}\label{sec:cite}

In order to justify the work the developers of \dealii{} put into this
software, we ask that papers using the library reference one of the
\dealii{} papers. This helps us justify the effort we put into it.

There are various ways to reference \dealii{}. To acknowledge the use of
the current version of the library, \textbf{please reference the present
  document}. For up to date information and a bibtex entry
see
\begin{center}
  \url{https://www.dealii.org/publications.html}
\end{center}

The original \dealii{} paper containing an overview of its
architecture is \cite{BangerthHartmannKanschat2007}. If you rely on
specific features of the library, please consider citing any of the
following:
\begin{multicols}{2}
  \vspace*{-36pt}
  \begin{itemize}
    \item For geometric multigrid: \cite{Kanschat2004,JanssenKanschat2011,ClevengerHeisterKanschatKronbichler2019};
    \item For distributed parallel computing: \cite{BangerthBursteddeHeisterKronbichler11};
    \item For $hp$-adaptivity: \cite{BangerthKayserHerold2007};
    \item For partition-of-unity (PUM) and enrichment methods of the
          finite element space: \cite{Davydov2016};
    \item For matrix-free and fast assembly techniques:
          \cite{KronbichlerKormann2012,KronbichlerKormann2019};
    \item For computations on lower-dimensional manifolds:
          \cite{DeSimoneHeltaiManigrasso2009};
    \item For curved geometry representations and manifolds:
          \cite{HeltaiBangerthKronbichlerMola2019};
          \vfill\null  \columnbreak
    \item For integration with CAD files and tools:
          \cite{HeltaiMola2015};
    \item For boundary element computations:
          \cite{GiulianiMolaHeltai-2018-a};
    \item For \texttt{LinearOperator} and \texttt{PackagedOperation} facilities:
          \cite{MaierBardelloniHeltai-2016-a,MaierBardelloniHeltai-2016-b};
    \item For uses of the \texttt{WorkStream} interface:
          \cite{TKB16};
    \item For uses of the \texttt{ParameterAcceptor} concept, the
          \texttt{MeshWorker::ScratchData} base class, and the
          \texttt{ParsedConvergenceTable} class:
          \cite{SartoriGiulianiBardelloni-2018-a};
    \item For uses of the particle functionality in \dealii{}:
          \cite{GLHPB18}.
          \vfill\null
  \end{itemize}
\end{multicols}

\dealii{} can interface with many other libraries:
\begin{multicols}{3}
  \begin{itemize}
    \item ADOL-C \cite{Griewank1996a,adol-c}
    \item ARPACK \cite{arpack}
    \item Assimp \cite{assimp}
    \item BLAS and LAPACK \cite{lapack}
    \item cuSOLVER \cite{cusolver}
    \item cuSPARSE \cite{cusparse}
    \item Gmsh \cite{geuzaine2009gmsh}
    \item GSL \cite{gsl2016}
    \item Ginkgo \cite{ginkgo-web-page}
    \item HDF5 \cite{hdf5}
    \item METIS \cite{karypis1998fast}
    \item MUMPS \cite{ADE00,MUMPS:1,MUMPS:2,mumps-web-page}
    \item muparser \cite{muparser-web-page}
    \item nanoflann \cite{nanoflann}%
          %
          % nanoflann has been deprecated and will be removed for 9.3. remove
          % this line in the 9.3 release paper
          %
    \item NetCDF \cite{rew1990netcdf}
          %
          % netcdf has been deprecated and will be removed for 9.3. remove
          % this line in the 9.3 release paper
          %
    \item OpenCASCADE \cite{opencascade-web-page}
    \item p4est \cite{p4est}
    \item PETSc \cite{petsc-user-ref,petsc-web-page}
    \item ROL \cite{ridzal2014rapid}
    \item ScaLAPACK \cite{slug}
    \item SLEPc \cite{Hernandez:2005:SSF}
    \item SUNDIALS \cite{sundials}
    \item SymEngine \cite{symengine-web-page}
    \item TBB \cite{Rei07}
    \item Trilinos \cite{trilinos,trilinos-web-page}
    \item UMFPACK \cite{umfpack}
  \end{itemize}
\end{multicols}
Please consider citing the appropriate references if you use
interfaces to these libraries. We note that the nanoflann and NetCDF
interfaces are now deprecated and will be removed in \dealii{} version
9.3.

The two previous releases of \dealii{} can be cited as
\cite{dealII91,dealII92}.

\section{Acknowledgments}

\dealii{} is a world-wide project with dozens of contributors around the
globe. Other than the authors of this paper, the following people
contributed code to this release:\\
%
% Uwe Koecher doesn't usually show up in the changelog, but
% we should make sure he's listed.
%
% 9.3: updated 5/08/2021 PM
Alexander Grayver,
Andrew Davis,
Arpit Babbar,
Bruno Blais,
Bruno Turcksin,
Ce Qin,
Christoph Kammer,
Conrad Clevenger,
Dominic Soldner,
Daniel Arndt,
Daniel Garcia-Sanchez,
Daniel Paukner,
David F. Castellanos,
David Wells,
Doug Shi-Dong,
David Schneider,
Elias Dejene,
Graham Harper,
Ignacio Tomas,
Ivan Fumagalli,
Jake Harmon,
Jean-Paul Pelteret,
Jiaqi Zhang,
Julian Roth,
Katharina Kormann,
Konrad Simon,
Krishnakumar Gopalakrishnan,
Laura Prieto Saavedra,
Lei Qiao,
Luca Heltai,
Magdalena Schreter,
Malhar Tidke,
Marc Fehling,
Marcus Calhoun-Lopez,
Martin Kronbichler ,
Mathias Anselmann,
Matthias Maier,
Maximilian Bergbauer,
Menno Fraters,
Michele Bucelli,
Nicola Giuliani,
Nicolas Barnafi,
Niklas Fehn,
Nils Much,
Olivier Guevremont,
Pasquale Africa,
Peter Munch,
Peter Westerbaan,
Phillip Mobley,
Pratik Nayak,
Praveen Chandrashekar,
Rene Gassmoeller,
Reza Rastak,
Richard Schussnig,
Sebastian Proell,
Simon Sticko,
Timo Heister,
Toni El Geitani Nehme,
Wenyu Lei,
Wolfgang Bangerth,
Zhou Lei

Their contributions are much appreciated!

\bigskip

\dealii{} and its developers are financially supported through a
variety of funding sources:

D.~Arndt and B.~Turcksin: Research sponsored by the Laboratory Directed Research and
Development Program of Oak Ridge National Laboratory, managed by UT-Battelle,
LLC, for the U. S. Department of Energy.

W.~Bangerth, T.~C.~Clevenger, and T.~Heister were partially
supported by the Computational Infrastructure
in Geodynamics initiative (CIG), through the National Science
Foundation under Award No.~EAR-1550901 and The
University of California -- Davis.

W.~Bangerth was also partially supported by award OAC-1835673 as part of the Cyberinfrastructure for Sustained Scientific Innovation (CSSI)
program, DMS-1821210,
and EAR-1925595.

%D.~Davydov was supported by the German Research Foundation (DFG), grant DA
%1664/2-1 and the Bayerisches Kompetenznetzwerk
%f\"ur Technisch-Wissenschaftliches Hoch- und H\"ochstleistungsrechnen
%(KONWIHR).

B.~Blais was partially supported by the National Science and Engineering Research Council of Canada(NSERC)  through the RGPIN-2020-04510 Discovery Grant

T.~C.~Clevenger was also partially supported EAR-1925575 and OAC-2015848.

A.~V.~Grayver was partially supported by the European Space Agency
Swarm DISC program.

Timo Heister was also partially supported by the National Science Foundation (NSF)
Award DMS-2028346, OAC-2015848, EAR-1925575, and by
Technical Data Analysis, Inc. through US Navy STTR Contract N68335-18-C-0011.

L.~Heltai was partially supported by the Italian Ministry of Instruction,
University and Research (MIUR), under the 2017 PRIN project NA-FROM-PDEs MIUR
PE1, ``Numerical Analysis for Full and Reduced Order Methods for the efficient
and accurate solution of complex systems governed by Partial Differential
Equations''.

M.~Kronbichler and P.~Munch were partially supported by the
Bayerisches Kompetenznetzwerk
f\"ur Technisch-Wissen\-schaft\-li\-ches Hoch- und H\"ochstleistungsrechnen
(KONWIHR) in the context of the projects
``Performance tuning of high-order discontinuous Galerkin solvers for
SuperMUC-NG'' and ``High-order matrix-free finite element implementations with
hybrid parallelization and improved data locality''.

M.~Maier was partially supported by ARO MURI Award No. W911NF-14-0247 and
NSF Award DMS-1912847.

D.~Wells was supported by the National Science Foundation (NSF) through Grant
DMS-1344962.

Z.~Wang was partially
supported by the National Science Foundation under award OAC-1835673.

The Interdisciplinary Center for Scientific Computing (IWR) at Heidelberg
University has provided hosting services for the \dealii{} web page.

The authors acknowledge the Texas Advanced Computing Center (TACC) at The
University of Texas at Austin for providing access to HPC resources that have
contributed to the research results reported within this paper.

Clemson University is acknowledged for generous allotment of compute time on
the Palmetto cluster.

\bibliography{paper}{}
\bibliographystyle{abbrv}

\end{document}