### 2024-02-28 2024-licensing_acknowledgment/relicensing.tex
add licensing acknowledgment for relicensing

\RequirePackage{fix-cm}
\documentclass[oneside, 12pt, letterpaper]{article}
\usepackage{fixltx2e}[2006/03/24]

\usepackage{geometry}
\geometry{left=1.5in, right=1.5in, top=1.25in, bottom=1.0in}
\usepackage{parskip}
\usepackage{graphicx}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[american]{babel}
\usepackage{enumitem}
\usepackage{amssymb}

% \usepackage[colorlinks=true,linkcolor=black,citecolor=black,urlcolor=black]{hyperref}
\usepackage{url}

\usepackage{microtype}
\microtypesetup{
  tracking=true,
  kerning=true
}
\microtypecontext{spacing=nonfrench}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\thispagestyle{empty}
\enlargethispage{3em}

\begin{center}
  {\bfseries deal.II Licensing Acknowledgment}
\end{center}

By signing below, I acknowledge and agree to the following:

{\bfseries (a)} "{\bfseries Contributions to the deal.II Project}" shall
mean any copyrightable material (including, but not limited to, software
source code, documentation source, configuration files, data, and images)
that I have made available to the deal.II project whether provided before,
on, or after the date hereof. This specifically includes prior
contributions to
\vspace{-0.5em}
\begin{itemize}[noitemsep]
  \item[--]
    the {\bfseries deal.II source code repository} currently hosted at
    \url{https://github.com/dealii/dealii} as part of the development
    history leading up to commit
    \texttt{65d590d6e8fb94ce1fd704d69cc1ed5b874857fb}, and the
  \item[--]
    the {\bfseries website repository} currently hosted at
    \url{https://github.com/dealii/old-website} as part of the development
    history leading up to commit
    \texttt{6e54856bae6ed17385989d7b6f08a659a5fa2d43},
\end{itemize}
\vspace{-0.5em}
but also any pending or future pull requests to these repositories.

{\bfseries (b)} The "{\bfseries Apache License 2.0 with LLVM-exception}"
and "{\bfseries GNU Lesser General Public License version 2.1 or later}"
are open source licenses whose full license texts are publicly available at
\vspace{-0.5em}
\begin{itemize}[noitemsep]
  \item[--] \url{https://spdx.org/licenses/Apache-2.0.html}
  \item[--] \url{https://spdx.org/licenses/LLVM-exception.html}
  \item[--] \url{https://spdx.org/licenses/LGPL-2.1-or-later.html}
\end{itemize}
\vspace{-0.5em}
The full license texts are also available in the file \texttt{LICENSE.md}
in the source code repository with commit
\texttt{65d590d6e8fb94ce1fd704d69cc1ed5b874857fb}.

{\bfseries (c)} I am legally entitled to grant the licenses granted herein.

{\bfseries (d)} I hereby license all {\bfseries Contributions to the
deal.II Project} under the terms of the {\bfseries Apache License 2.0 with
LLVM-exception}. Furthermore, I affirm that my {\bfseries Contributions to
the deal.II Project} which have been previously licensed under the terms of
the {\bfseries GNU Lesser General Public License v2.1 or later} will
continue to also be available under these terms.

{\bfseries (e)} I choose to {\bfseries $\square$\, include /
$\square$\, exclude} my contributions to the {\bfseries Code Gallery
repository} currently hosted at
\url{https://github.com/dealii/code-gallery} into/from this license grant.
If no selection is made, then ``exclude'' applies.
\vspace{0.5em}

Full name:
\vspace{0.5em}

E-Mail address(es):
\vspace{2.75em}

Date:
\vspace{0.5em}

Signature:

\end{document}